---
title: "Anlysis of empirical victim-exploiter bipartite networks"
---


This script reads a collection of empirical victim-exploiter bipartite networks, removes isolated species, and computes a set of structural metrics at both the network and species levels. The dataset of empirical networks metrics it produces is then used in **2_simulation_empirical_networks.Rmd** and in **3_metrics_simulated_networks.Rmd**.

For each network, it quantifies species richness, connectance, modularity, nestedness (NODF), degree distributions, complexity indices, and matrix entropy (not included in the article), alongside species-level degree and betweenness centrality. 

It then computes total indirect interaction effects from the bipartite adjacency matrix, decomposing these effects into incoming (INN) and outgoing (OUT) components, and summarizes their magnitude, variability, and evenness using (normalized and non-normalized) Shannon entropy. 

All metrics are stored in a unified data frame and saved, and visualized through a series of exploratory plots.


```{r}
library(poweRlaw)
library(bipartite)
library(vegan)
library(igraph)
library(tidyverse)
library(ggplot2)
library(reshape2)
library(tidyr)
library(dplyr)
library(ggpubr)
library(viridis)
library(grid)
library(bipartite) # to use the networklevel() function to calculate modularity and nestedness

source(here::here("code/functions/Functions.R"))
source(here::here("code/functions/compute_net_effects.R"))
source(here::here("code/functions/compute_metrics_networks.R"))


```


For plotting:

```{r}
my_theme<-theme_classic()+theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)))

```



### Read and process empirical networks

```{r}

# Path to empirical network files

path <-"../data/empirical_networks"
files_path <- list.files(path = path, pattern = ".txt", full.names = TRUE)


# read metadata

table_networks<-read.csv("../data/empirical_networks/antagonistic_networks.csv", sep=",", header = TRUE)

# optain list of names

names_networks<- unname(sapply(table_networks[, "network.file"], paste, ".txt", sep=""))


# create list of network .csv files

list_networks <- list()

for (i in 1:length(files_path)) {

  
  list_networks[[i]] <-read.csv(files_path[i], sep="", header = FALSE)
  
}


# check we read them all

if(length(list_networks) == length(files_path)){
  
  print("all good")
}else{
  print("problem")
}


# add list of network .csv files to the metadata table

table_networks$list_networks <- list_networks
table_networks$names_network_files <-names_networks

table_networks <- table_networks %>% 
    rename("nvictim" = "nvitim",
           "connect" = "connectance") %>% 
  select(-nvictim, 
         -nexpl,
         -connect)

head(table_networks)

```


Remove isolated species

```{r}

table_networks$list_networks <- lapply(table_networks$list_networks, remove_isolated_species)

```



## Compute Network metrics

```{r}

# function to compute connectance

compute_connectance <- function(g) {
  # Ensure the graph is bipartite
  if (!is_bipartite(g)) {
    stop("The graph is not bipartite. Ensure 'type' attribute is correctly set.")
  }
  
  # Extract node types
  type <- V(g)$type
  
  # Number of species at each level
  T <- sum(type)       # Number of nodes in the top level (TRUE in 'type')
  B <- sum(!type)      # Number of nodes in the bottom level (FALSE in 'type')
  
  # Number of observed edges (L)
  L <- ecount(g)       # Total number of edges in the bipartite graph
  
  # Maximum possible interactions (m)
  m <- T * B           # Product of species in the two levels
  
  print(paste("T =", T, "B =", B, "L =", L, "m =", m))
  if (m == 0 || is.na(m)) {
    stop("Invalid maximum possible interactions (m).")
  }
  
  # Connectance formula
  connectance <- L / m
  return(connectance)
}

compute_modularity <- function(g) {
  # Compute modularity using different methods
  mod_louvain <- cluster_louvain(g)$modularity
  mod_walktrap <- modularity(cluster_walktrap(g))
  mod_fast_greedy <- modularity(cluster_fast_greedy(g))
  mod_leading_eigen <- cluster_leading_eigen(g)$modularity
  
  # Compute the mean modularity
  mean_modularity <- mean(c(mod_louvain, mod_walktrap, mod_fast_greedy, mod_leading_eigen), na.rm = TRUE)
  
  return(mean_modularity)
}
```


SD entropy computed as:


$$ J = - \frac{1}{ln(k)} * \sum_{i = 1}^{k} s_i * ln(s_i)  $$


```{r}

# network level

table_networks$nvictim <- vector("numeric", length = nrow(table_networks))
table_networks$nexpl <- vector("numeric", length = nrow(table_networks))
table_networks$nspp <- vector("numeric", length = nrow(table_networks))
table_networks$connectance <- vector("numeric", length = nrow(table_networks))
table_networks$modularity <- vector("numeric", length = nrow(table_networks))
table_networks$NODF <- vector("numeric", length = nrow(table_networks))
table_networks$alpha <- vector("numeric", length = nrow(table_networks))
table_networks$complexity_index <- numeric(length(table_networks$list_networks))
table_networks$entropy <- numeric(length(table_networks$list_networks))


# spp level

table_networks$betw <- vector("list", nrow(table_networks))
table_networks$betw_v <- vector("list", nrow(table_networks))
table_networks$betw_e <- vector("list", nrow(table_networks))
table_networks$degree <- vector("list", nrow(table_networks))
table_networks$degree_v <- vector("list", nrow(table_networks))
table_networks$degree_e <- vector("list", nrow(table_networks))

# Loop through each network in the list

counter <- 0

for (i in 1:length(table_networks$list_networks)) {
  
  counter <- counter + 1  
  
  # Read data
  network <- table_networks$list_networks[[i]]
  
  # Convert matrix to igraph object
  g <- graph_from_biadjacency_matrix(network)
  
  # Compute metrics
  table_networks$nvictim[i] <- nrow(network)
  table_networks$nexpl[i] <- ncol(network)
  table_networks$nspp[i] <- sum(dim(network))
  table_networks$connectance[i] <- unname(compute_connectance(g))
  table_networks$modularity[i] <- unname(compute_modularity(g))
  table_networks$NODF[i] <- unname(nestednodf(network)$statistic["NODF"])
  graph <- igraph::graph_from_incidence_matrix(network)
  degrees <- igraph::degree(graph)
  table_networks$alpha[i] <- compute_alpha_pl(degrees)
  table_networks$complexity_index[i] <- sqrt(unname(networklevel(network, index = "connectance")) * sum(dim(network)))
  table_networks$entropy[i] <- compute_SVD_entropy(network)
  
  table_networks$betw[[i]] <- compute_centrality(binmatrix = unname(as.matrix(table_networks$list_networks[[i]])),
                                                 n_v = nrow(network), n_e = ncol(network))$betw
  
  table_networks$betw_v[[i]] <- compute_centrality(binmatrix = unname(as.matrix(table_networks$list_networks[[i]])),
                                                   n_v = nrow(network), n_e = ncol(network))$betw_v
  
  table_networks$betw_e[[i]] <- compute_centrality(binmatrix = unname(as.matrix(table_networks$list_networks[[i]])),
                                                   n_v = nrow(network), n_e = ncol(network))$betw_e
  
  table_networks$degree[[i]] <- compute_centrality(binmatrix = unname(as.matrix(table_networks$list_networks[[i]])),
                                                   n_v = nrow(network), n_e = ncol(network))$degree
  
  table_networks$degree_v[[i]] <- compute_centrality(binmatrix = unname(as.matrix(table_networks$list_networks[[i]])),
                                                     n_v = nrow(network), n_e = ncol(network))$degree_v
  
  table_networks$degree_e[[i]] <- compute_centrality(binmatrix = unname(as.matrix(table_networks$list_networks[[i]])),
                                                     n_v = nrow(network), n_e = ncol(network))$degree_e
  
  
  # Print progress
  cat("Iteration ", counter, "/", length(table_networks$list_networks), "\n")
}

head(table_networks)

```




```{r}
# Names to exclude
exclude_files <- c("lebsko_complex.txt", "wichels.txt")

# Filter out rows with these names
table_networks <- table_networks %>%
  filter(!names_network_files %in% exclude_files)
```


## Compute net effects

nf1 = INN effects

nf2 = OUT effects

```{r}


# community-level
table_networks$T_matrix <- vector(mode = "list", length = nrow(table_networks))
table_networks$T_mean <- numeric(nrow(table_networks))
table_networks$T_sd <- numeric(nrow(table_networks))


# cummulative effects

### nf1

table_networks$T_nf1_s_vec <- vector(mode = "list", length = nrow(table_networks))
table_networks$T_nf1_sum_vec <- vector(mode = "list", length = nrow(table_networks))
table_networks$T_nf1_mean_vec <- vector(mode = "list", length = nrow(table_networks))
table_networks$T_nonorm_nf1_s_vec <- vector(mode = "list", length = nrow(table_networks))

table_networks$T_nf1_sum <- numeric(nrow(table_networks))
table_networks$T_nf1_mean <- numeric(nrow(table_networks))
table_networks$T_nf1_sd <- numeric(nrow(table_networks))
table_networks$T_nf1_s <- numeric(nrow(table_networks)) # shannon index
table_networks$T_nonorm_nf1_s <- numeric(nrow(table_networks))



### nf2

table_networks$T_nf2_s_vec <- vector(mode = "list", length = nrow(table_networks))
table_networks$T_nf2_sum_vec <- vector(mode = "list", length = nrow(table_networks))
table_networks$T_nf2_mean_vec <- vector(mode = "list", length = nrow(table_networks))
table_networks$T_nonorm_nf2_s_vec <- vector(mode = "list", length = nrow(table_networks))


table_networks$T_nf2_sum <- numeric(nrow(table_networks))
table_networks$T_nf2_mean <- numeric(nrow(table_networks))
table_networks$T_nf2_sd <- numeric(nrow(table_networks))
table_networks$T_nf2_s <- numeric(nrow(table_networks)) # shannon index
table_networks$T_nonorm_nf2_s <- numeric(nrow(table_networks))




for (i in 1:nrow(table_networks)) {
  mat <- unname(as.matrix(table_networks$list_networks[[i]]))
  
  # Calculate total effects matrix
  T <- compute_total_effects(mat)
  
  # Store total effects matrix
  table_networks$T_matrix[[i]] <- T
  
  # Compute mean and standard deviation of total effects matrix
  table_networks$T_mean[i] <- mean(T)
  table_networks$T_sd[i] <- sd(T)
  
  # Compute cummulative effects 
  
  table_networks$T_nf1_sum_vec[[i]] <- rowSums(T)
  table_networks$T_nf2_sum_vec[[i]] <- colSums(T)
  
  table_networks$T_nf1_sum[i] <- mean(rowSums(T))
  table_networks$T_nf2_sum[i] <- mean(colSums(T))
  
  
  table_networks$T_nf1_mean_vec[[i]] <- rowMeans(T)
  table_networks$T_nf2_mean_vec[[i]] <- colMeans(T)
  
  table_networks$T_nf1_mean[i] <- mean(rowMeans(T))
  table_networks$T_nf2_mean[i] <- mean(colMeans(T))
  
  table_networks$T_nf1_sd[i] <- sd(rowMeans(T))
  table_networks$T_nf2_sd[i] <- sd(colMeans(T))
  
  
  # Compute shannon index
  
  T_nf1_s_vec <- apply(T, 1, compute_shannon)
  T_nf2_s_vec <- apply(T, 2, compute_shannon)
  T_nonorm_nf1_s_vec <- apply(T, 1, compute_shannon_nonorm)
  T_nonorm_nf2_s_vec <- apply(T, 2, compute_shannon_nonorm)
  
  # Store shannon index nf1 and nf2 vectors
  table_networks$T_nf1_s_vec[[i]] <- T_nf1_s_vec
  table_networks$T_nf2_s_vec[[i]] <- T_nf2_s_vec
  table_networks$T_nonorm_nf1_s_vec[[i]] <- T_nonorm_nf1_s_vec
  table_networks$T_nonorm_nf2_s_vec[[i]] <- T_nonorm_nf2_s_vec
  
  
  
  table_networks$T_nf1_s[i] <- mean(T_nf1_s_vec)
  table_networks$T_nf2_s[i] <- mean(T_nf2_s_vec)
  table_networks$T_nonorm_nf1_s[i] <- mean(T_nonorm_nf1_s_vec)
  table_networks$T_nonorm_nf2_s[i] <- mean(T_nonorm_nf2_s_vec)
  
}


head(table_networks)
```

Save table

```{r}
saveRDS(table_networks, here::here("output/df_network_metrics.rds"))
```


## Plot metrics


```{r}

# Plot modularity against nestedness
plot_mod_nest <- ggplot(table_networks, aes(x = modularity, y = NODF)) +
  geom_point(alpha = 0.3, size = 2) +
  labs(x = "Modularity", y =expression(paste("Nestedness ",italic((NODF))))) +
  my_theme

# Plot connectance against NSPP
plot_conn_nspp <- ggplot(table_networks, aes(x = connectance, y = nspp)) +
  geom_point(alpha = 0.3, size = 2) +
  labs(x = "Connectance", y = "Species richness") +
  my_theme

# Plot nexpl against nvictim
plot_nexpl_nvictim <- ggplot(table_networks, aes(x = nexpl, y = nvictim)) +
  geom_point(alpha = 0.3, size = 2) +
  labs(x = expression(paste(italic(n)," exploiters")), y = expression(paste(italic(n)," victims"))) +
  my_theme

# Plot distribution of alpha values
plot_alpha_dist <- ggplot(table_networks, aes(x = alpha)) +
  geom_histogram(binwidth = 0.1, fill = "black", color = "black", alpha = 0.3) +
  labs(x = expression(alpha)) +
  my_theme

# Plot distribution of complexity index
plot_compidx_dist <- ggplot(table_networks, aes(x = complexity_index)) +
  geom_histogram(binwidth = 0.1, fill = "black", color = "black", alpha = 0.3) +
  labs(x = expression(sqrt(C * S))) +
  my_theme


# Plot distribution of entropy with a vertical line at x = 1
#plot_entropy_dist <- ggplot(table_networks, aes(x = entropy)) +
 # geom_histogram(binwidth = 0.1, fill = "black", color = "black", alpha = 0.3) +
 # geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
 # labs(x = expression(paste(italic(SVD), " entropy"))) +
 # my_theme

# Plot entropy_d against complexity index
#plot_entropy_compidx <- ggplot(table_networks, aes(x = entropy, y = complexity_index)) +
 # geom_point(alpha = 0.3, size = 2) +
 # labs(x = expression(paste(italic(SVD), " entropy")), y = expression(sqrt(C * S))) +
 # my_theme

# Combine plots
combined_plots <- ggarrange(
  plot_mod_nest + labs(title = NULL),
  plot_conn_nspp + labs(title = NULL),
  plot_nexpl_nvictim + labs(title = NULL),
  plot_alpha_dist + labs(title = NULL),
  plot_compidx_dist + labs(title = NULL),
  ncol = 2, nrow = 3,
  
  labels = LETTERS[1:5]
)

combined_plots

# Save the combined plot
# ggsave(here::here("figures/supplmat/combined_plots.png"), combined_plots, width = 7, height = 9)

```




# Net effects on species ~ structure



```{r}
p_nev_h <- ggplot(table_networks, aes(x = nexpl, y = nvictim)) +  
    geom_point(aes(color = T_nf1_s, size = nspp), alpha = 0.6) +  
    labs(x = expression(paste(italic(n), " exploiters")), y = expression(paste(italic(n), " victims"))) +
    scale_color_viridis_c(name = expression(italic((H[b]))),
                          option = "turbo") +
    my_theme


p_c_h <- ggplot(table_networks, aes(x = connectance, y = T_nf1_s)) +  
    geom_point(aes(size = nspp), alpha = 0.4) +  
    labs(x = "Connectance", y = expression(paste("Biotic effects evennness ",italic((H[b]))))) +
  #  scale_color_viridis_c(name = "Mean effect inn") +
    my_theme

ggarrange(
  
  p_nev_h,
  p_c_h,
  
  ncol = 2,
  nrow = 1,
  
  common.legend = TRUE,
  
  labels = LETTERS[1:2]
)

#ggsave(here::here("figures/supplmat/nve_c_h.png"), height = 7, width = 10)

```



## Plots


### Normalized shannon


```{r}
  
  # Plot modularity against nestedness with T_mean and T_nf1_mean
  p_mod_nest_effects <-   ggplot(table_networks, aes(x = modularity, y = NODF)) +  
    geom_point(aes(color = T_nf1_s, size = nspp), alpha = 0.7) +  
    labs(x = "Modularity", y = expression(paste("Nestedness ", italic((NODF)))),
           size = expression(italic(S))) +
        scale_color_viridis_c(name = expression(italic((H[b]))), option = "turbo") +
    my_theme
  
  
  # Plot distribution of complexity index with T_mean and T_nf1_mean as vertical lines
  plot_complexity_effects <- ggplot(table_networks, aes(x = complexity_index, y = T_nf1_s)) +
    geom_point( alpha = 0.5, size = 3) +
    labs(x = expression(sqrt(C * S)),
         y =expression(paste("Biotic effects evennness ",italic((H))))) +
    my_theme
  

  
  
  
  df_centrality <- data.frame("betw" = unlist(table_networks$betw),
                              "degree" = unlist(table_networks$degree),
                              "T_nf1_s_vec" = unlist(table_networks$T_nf1_s_vec))
  
  
  
  plot_degree_effects <- ggplot(df_centrality, aes(x = sqrt(degree), y = T_nf1_s_vec)) +
    geom_point( alpha = 0.3) +
    labs(x = "Species degree",
         y = expression(paste("Biotic effects evennness ",italic((H))))) +
    scale_x_continuous(labels = back_transform_sqrt)+
    my_theme
  
  
  
  fig1 <- ggarrange(
    
    plot_complexity_effects,
    p_mod_nest_effects,
    plot_degree_effects,
    
    
    ncol = 3,
    nrow = 1,
    
    common.legend = TRUE,
    
    labels = c("D", "E", "F")
    
    
  )
  
  fig1
```

Save plots as objects to combine them afterwards in script **simulation_simulated_networks.Rmd** and form figure 4.

```{r}
#ggsave(here::here("figures/fig1.png"), height = 5, width = 12)
saveRDS(plot_complexity_effects,here::here("figures/plot_complexity_effects.rds"))
saveRDS(p_mod_nest_effects,here::here("figures/p_mod_nest_effects.rds"))
saveRDS(plot_degree_effects,here::here("figures/plot_degree_effects.rds"))
```




### Non-normalized shannon

```{r}

  # Plot modularity against nestedness with T_mean and T_nf1_mean
  p_mod_nest_effects_nonorm <-   ggplot(table_networks, aes(x = modularity, y = NODF)) +  
    geom_point(aes(color = T_nonorm_nf1_s, size = nspp), alpha = 0.7) +  
    labs(x = "Modularity", y = expression(paste("Nestedness ", italic((NODF)))),
           size = expression(italic(S))) +
        scale_color_viridis_c(name = expression(italic((H))), option = "turbo") +
    my_theme
  
  
  # Plot distribution of complexity index with T_mean and T_nf1_mean as vertical lines
  plot_complexity_effects_nonorm <- ggplot(table_networks, aes(x = complexity_index, y = T_nonorm_nf1_s)) +
    geom_point( alpha = 0.5, size = 3) +
    labs(x = expression(sqrt(C * S)),
         y =expression(paste("Biotic effects evennness ",italic((H))))) +
    my_theme
  

  
  
  
  df_centrality <- data.frame("betw" = unlist(table_networks$betw),
                              "degree" = unlist(table_networks$degree),
                              "T_nonorm_nf1_s_vec" = unlist(table_networks$T_nonorm_nf1_s_vec))
  
  
  
  plot_degree_effects_nonorm <- ggplot(df_centrality, aes(x = sqrt(degree), y = T_nonorm_nf1_s_vec)) +
    geom_point( alpha = 0.3) +
    labs(x = "Species degree",
         y = expression(paste("Biotic effects evennness ",italic((H))))) +
    scale_x_continuous(labels = back_transform_sqrt)+
    my_theme
  
  
  
  
  
  plot_complexity_effects_nonorm
  p_mod_nest_effects_nonorm
  plot_degree_effects_nonorm
  
  
    
  fig_appendix_hraw <- ggarrange(
    
    plot_complexity_effects_nonorm,
    p_mod_nest_effects_nonorm,
    plot_degree_effects_nonorm,
    
    
    ncol = 3,
    nrow = 1,
    
    common.legend = TRUE,
    
    labels = c("A", "B", "C")
    
    
  )
  
  fig_appendix_hraw
  

```


#### comparison normalized vs non-normalized shannon

```{r}
# Combine normalized and raw values into a single data frame
df_all_values <- data.frame(
  H_norm = unlist(table_networks$T_nf1_s_vec),
  H_raw = unlist(table_networks$T_nonorm_nf1_s_vec)
)

# Compute the Pearson correlation
correlation_all <- cor(df_all_values$H_norm, df_all_values$H_raw, use = "complete.obs")

# Scatter plot of normalized vs raw values
ggplot(df_all_values, aes(x = H_norm, y = H_raw)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(
    x = expression(paste("Normalized biotic effects evennness ",italic((H)))),
    y = expression(paste("Raw biotic effects evennness ",italic((H)))),
    title = paste("Pearson correlation =", round(correlation_all, 2))
  ) +
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0.5) # Center the title
  )

#ggsave(here::here("figures/figS_hcor.png"), height = 5, width = 4)

```
