---
title: "Coevolution simulation using empirical networks"
---

This script runs coevolutionary simulations on each empirical network and generates figures illustrating the relationships between network structural properties and measures of coevolution.

The simulation includes an option to introduce abiotic perturbations by shifting species' environmental optima. These perturbation scenarios were not used in the article **Trapped in the web: network architectures spread coevolution and shape adaptation**. The simulation function includes an argument that controls whether such perturbations are applied and is currently set to exclude them.


```{r}
library(tidyverse)
library(ggplot2)
library(reshape2)
library(tidyr)
library(dplyr)
library(ggpubr)
library(viridis)
library(grid)
library(bipartite) # to use the networklevel() function to calculate modularity and nestedness
library(igraph)

# Custom functions

source("Functions.R")
source("load_computation.R")
```



## prepare scenarios' dataframe

Read empirical networks table

```{r}

df_network_metrics <- readRDS(here::here("output/df_network_metrics.rds"))

df_network_metrics <- df_network_metrics %>%  # rename column with network names
  rename(id_network = network.file)

head(df_network_metrics)
```



*Make dataframe*:


```{r}
nrepl <- 20

# Create combinations of scenarios based on network and perturbation
df_scenarios <- expand.grid(id_network = df_network_metrics$id_network)

# Generate 10 replicates for each scenario
df_replicates <- df_scenarios %>%
  mutate(scenario = row_number()) %>%
  tidyr::crossing(replicate = 1:nrepl) %>%
  arrange(scenario, replicate) 

# Print or use df_replicates as needed
head(df_replicates)

```



## set storing objects

```{r}

df_replicates <- df_replicates %>% 
  mutate(
  
  
# trait values at equilibrium
    
traits_eq = vector("list", nrow(df_replicates)),
traits_eq_push1 = vector("list", nrow(df_replicates)),
traits_eq_push2 = vector("list", nrow(df_replicates)),
teta = vector("list", nrow(df_replicates)),



# evolutionary rate

comm_evolr = vector("list", nrow(df_replicates)),
v_evolr = vector("list", nrow(df_replicates)),
e_evolr = vector("list", nrow(df_replicates)),


# mean fluctuation in trait matching

comm_fcij = vector("list", nrow(df_replicates)),
v_fcij = vector("list", nrow(df_replicates)),
e_fcij = vector("list", nrow(df_replicates)),
    

# mean fitness load

comm_load_total = vector("list", nrow(df_replicates)),
comm_load_biot = vector("list", nrow(df_replicates)),
comm_load_abiot = vector("list", nrow(df_replicates)),
comm_load_network = vector("list", nrow(df_replicates)),

v_load_total = vector("list", nrow(df_replicates)),
v_load_biot = vector("list", nrow(df_replicates)),
v_load_abiot = vector("list", nrow(df_replicates)),
v_load_network = vector("list", nrow(df_replicates)),

e_load_total = vector("list", nrow(df_replicates)),
e_load_biot = vector("list", nrow(df_replicates)),
e_load_abiot = vector("list", nrow(df_replicates)),
e_load_network = vector("list", nrow(df_replicates)),


)


df_replicates$id_network <- as.character(df_replicates$id_network)
df_replicates <- as.data.frame(df_replicates)

head(df_replicates)

```



## create dataframes for perturbation scenarios

```{r}
df_replicates_pert1 = df_replicates 
df_replicates_pert1$time_to_equilibrium <- numeric(nrow(df_replicates_pert1))
df_replicates_pert2 = df_replicates 
df_replicates_pert2$time_to_equilibrium <- numeric(nrow(df_replicates_pert2))
```


## set fixed parameters

```{r}

nsims <- 300
her <- 0.22 # heritability
e <- 0.5 # # e (epsilon trait distance)
b <- 10 # b (b expoent exploiter preference)


# Set values of biotic selection

bio_sel_v<-0.16
bio_sel_e<-0.16

ksi_hosts<-1-bio_sel_v
ksi_exploiters<- 1-bio_sel_e


# simulate perturbations?

simulate_perturbation1 = "no"
simulate_perturbation2 = "no"

# number of push steps

T_push <- 15
delta_teta <- 0.2
max_attempts <- 30 # max. number of attempts to find an equilibrium

```



# Simulation


```{r}
set.seed(1000)

for (sc in 1:nrow(df_replicates)) {
  print(paste(sprintf("Right now I am doing iteration %s", sc), "of", nrow(df_replicates)))
  
  # Get the network
  binary.matrix <- unname(as.matrix(df_network_metrics[df_network_metrics$id_network == df_replicates[sc, "id_network"], "list_networks"][[1]]))
  bigmat.F <- create_F(binary.matrix)
  
  # Set the number of species
  n.host <- nrow(binary.matrix)
  n.exploiter <- ncol(binary.matrix)
  nsp <- n.host + n.exploiter
  
  # Initialize simulation parameters
  ksi <- c(rep(ksi_hosts, n.host), rep(ksi_exploiters, n.exploiter))  # Strength of abiotic selection
  max_stable_timesteps <- 20  # Equilibrium condition
  equilibrium_reached <- FALSE  # Flag for equilibrium
  
attempt <- 0 # attempts for finding an equilibrium
  ### Equilibrium in Normal Conditions ###
  while (!equilibrium_reached && attempt < max_attempts) {
    attempt <- attempt + 1
    # Initialize matrices for traits and changes at the start of each attempt
    trait <- matrix(0, nrow = nsims, ncol = nsp)  # Trait values for each timestep
    z.change <- matrix(0, nrow = nsims, ncol = nsp)  # Trait changes for each timestep
    trait[1, ] <- rnorm(nsp, mean = 0, sd = 0.1)  # Initial random trait values
    teta <- trait[1, ]  # Environmental optima
    cummulative.matching.change <- matrix(0, nrow = n.host, ncol = n.exploiter)
    df_replicates$teta[[sc]] <- teta  # Save initial teta
    stable_timesteps <- 0  # Reset stable timesteps count
    stable_timesteps <- 0
    
    for (sim in 1:(nsims - 1)) {
      
      # Create matrix of differences in trait values for those species that interact
      a.matrix <- create_a_matrix(trait_vec = trait[sim, ], Fmat = bigmat.F)
      
      #Calculate env.sel for each sp
      E <- ksi * (teta - trait[sim, ])
      
      # Calculate interaction selection for victims & store it in vector Rh, and for exploiters in vetor Rp
      Rh <- calc_Rh_corrected(a_matrix = a.matrix, big_mat = bigmat.F, KSI.b = bio_sel_v, n_host = n.host, e_threshold = e)
      Rp <- calc_Rp_corrected(a_matrix = a.matrix, big_mat = bigmat.F, KSI.b = bio_sel_e, B = b, n_host = n.host, n_exploiter = n.exploiter, nsp = nsp)
      R <- c(Rh, Rp)
      
      # Compute Trait evolution
      z.change[sim, ] <- her * (R + E)
      trait[sim + 1, ] <- trait[sim, ] + z.change[sim, ]
      
      # Update a.matrix
      dif.traits2 <- outer(trait[sim+1,], trait[sim+1,], FUN= "-") 
      a.matrixsim2 <- bigmat.F*dif.traits2 
      
      # Calculate change in trait matching with time
      matching.change.matrix <- a.matrixsim2[1:n.host, (n.host+1):nsp] -a.matrix[1:n.host, (n.host+1):nsp]
      cummulative.matching.change <- cummulative.matching.change + abs(matching.change.matrix)
      
      # Check for equilibrium
      if (sim >= max_stable_timesteps && isTRUE(all.equal(trait[sim, ], trait[sim - max_stable_timesteps + 1, ], tolerance = 1e-10))) {
        stable_timesteps <- stable_timesteps + 1
        if (stable_timesteps >= max_stable_timesteps) {
          equilibrium_reached <- TRUE
          message("Equilibrium reached at timestep ", sim)
          df_replicates$traits_eq[[sc]] <- trait[sim, ]
          break
        }
      } else {
        stable_timesteps <- 0
      }
    }
    if (!equilibrium_reached) message("Restarting simulation: equilibrium not reached.")
  }
  
  if (!equilibrium_reached) {
  message("normal scenario - Skipping to the next iteration after ", max_attempts, " attempts.")
  next
}
  
  
  
  ################ store results
  
  # Evolutionary rate
  df_replicates$comm_evolr[[sc]] <- colSums(abs(z.change))/nsims
  df_replicates$v_evolr[[sc]] <- colSums(abs(z.change[,1:n.host]))/nsims
  df_replicates$e_evolr[[sc]]  <- colSums(abs(z.change[,(n.host+1):nsp]))/nsims
  
  # mean fluctuation in trait matching
  df_replicates$v_fcij[[sc]] <-rowSums(cummulative.matching.change)
  df_replicates$e_fcij[[sc]] <- colSums(cummulative.matching.change)
  df_replicates$comm_fcij[[sc]] <- c(rowSums(cummulative.matching.change), colSums(cummulative.matching.change))
  
  # fitness load
  
# --- Victims
simv_load_total <- numeric(n.host)
simv_load_abiot <- numeric(n.host)
simv_load_biot  <- numeric(n.host)

for (v in 1:n.host) {
  # Filter partners by interaction matrix
  interacting_exploiters <- which(binary.matrix[v, ] == 1)  # Get indices of interacting exploiters
  x_partners <- trait[sim, n.host + interacting_exploiters]  # Traits of interacting exploiters
  
  fitness_load <- fitness_load_victims(
    x_s = trait[sim, v],
    x_0 = teta[v],
    x_partners = x_partners,  # Only interacting exploiters
    epsilon = e  # Pass interaction threshold
  )
  
  simv_load_total[v] <- fitness_load$load
  simv_load_abiot[v] <- fitness_load$abiotic_L
  simv_load_biot[v] <- fitness_load$biotic_L
}

df_replicates$v_load_total[[sc]] <- simv_load_total
df_replicates$v_load_abiot[[sc]] <- simv_load_abiot
df_replicates$v_load_biot[[sc]] <- simv_load_biot

# --- Exploiters
sime_load_total   <- numeric(n.exploiter)
sime_load_abiot   <- numeric(n.exploiter)
sime_load_biot    <- numeric(n.exploiter)
sime_load_network <- numeric(n.exploiter)

for (p in (n.host + 1):nsp) {
  col_in_bin <- p - n.host                 
  hosts <- which(binary.matrix[, col_in_bin] == 1)

  if (length(hosts) == 0) {
    # no partners: only abiotic matters
    abi <- (trait[sim, p] - teta[p])^2
    sime_load_total[col_in_bin]   <- abi
    sime_load_abiot[col_in_bin]   <- abi
    sime_load_biot[col_in_bin]    <- 0
    sime_load_network[col_in_bin] <- 0
    next
  }

  x_partners <- trait[sim, hosts]         

  fl <- fitness_load_exploiters(
    x_s        = trait[sim, p],
    x_0        = teta[p],
    x_partners = x_partners
  )

  sime_load_total[col_in_bin]   <- fl$load
  sime_load_abiot[col_in_bin]   <- fl$abiotic_L
  sime_load_biot[col_in_bin]    <- fl$biotic_L
  sime_load_network[col_in_bin] <- fl$network_L
}


df_replicates$e_load_total[[sc]] <- sime_load_total
df_replicates$e_load_abiot[[sc]] <- sime_load_abiot
df_replicates$e_load_biot[[sc]] <- sime_load_biot
df_replicates$e_load_network[[sc]] <- sime_load_network
  
  
  #############
  teta_eq_unperturbed <- teta
  #############
  
  
  ########################
  # PUSH PERTURBATION
  ########################
  
  ### Perturbation Scenarios ###
  
  if (simulate_perturbation1 == "yes") {
    
    # Initialize results
  mean_time_to_equilibrium <- 0
  time_to_equilibrium_steps <- numeric(T_push)
  final_traits <- NULL
  
  # Initialize theta and traits from previous equilibrium
  teta <- teta_eq_unperturbed
  trait <- matrix(0, nrow = nsims, ncol = nsp)  # Store trait evolution
  z.change <- matrix(0, nrow = nsims, ncol = nsp)  # Store trait changes
  trait[1, ] <- df_replicates$traits_eq[[sc]]  # Traits from previous equilibrium
  cummulative.matching.change <- matrix(0, nrow=n.host, ncol=n.exploiter)
  
  
  for (step in 1:T_push) {
    
    # Gradual change in theta for hosts
    teta[1:n.host] <- teta[1:n.host] + delta_teta
    
    # Reset for equilibrium finding
    equilibrium_reached <- FALSE
    step_time <- 0
    stable_timesteps <- 0
    
    attempt <- 0 # attempts for finding an equilibrium
    while (!equilibrium_reached && attempt < max_attempts) {
      attempt <- attempt + 1
      
      # Reset the trait dynamics for this step
      z.change <- matrix(0, nrow = nsims, ncol = nsp)
      cummulative.matching.change <- matrix(0, nrow = n.host, ncol = n.exploiter)
      
      
      for (sim in 1:(nsims - 1)) {
        
        
        # Compute trait dynamics
        
        # Create matrix of differences in trait values for those species that interact
        a.matrix <- create_a_matrix(trait_vec = trait[sim, ], Fmat = bigmat.F)
        E <- ksi * (teta - trait[sim, ])
        
        # Calculate interaction selection for victims & store it in vector Rh, and for exploiters in vetor Rp
        Rh <- calc_Rh_corrected(a_matrix = a.matrix, big_mat = bigmat.F, KSI.b = bio_sel_v, n_host = n.host, e_threshold = e)
        Rp <- calc_Rp_corrected(a_matrix = a.matrix, big_mat = bigmat.F, KSI.b = bio_sel_e, B = b, n_host = n.host, n_exploiter = n.exploiter, nsp = nsp)
        R <- c(Rh, Rp)
        
        # Compute Trait evolution
        z.change[sim, ] <- her * (R + E)
        trait[sim + 1, ] <- trait[sim, ] + z.change[sim, ]
        
        # Update a.matrix
      dif.traits2 <- outer(trait[sim+1,], trait[sim+1,], FUN= "-") 
      a.matrixsim2 <- bigmat.F*dif.traits2 
      
      # Calculate change in trait matching with time
      matching.change.matrix <- a.matrixsim2[1:n.host, (n.host+1):nsp] -a.matrix[1:n.host, (n.host+1):nsp]
        cummulative.matching.change <- cummulative.matching.change + abs(matching.change.matrix)
        
        
        # Check for equilibrium
        if (sim >= max_stable_timesteps && isTRUE(all.equal(trait[sim, ], trait[sim - max_stable_timesteps + 1, ], tolerance = 1e-10))) {
          stable_timesteps <- stable_timesteps + 1
          if (stable_timesteps >= max_stable_timesteps) {
            equilibrium_reached <- TRUE
            step_time <- sim
            final_traits <- trait[sim, ]
            break
          }
        } else {
          stable_timesteps <- 0
        }
        
      }
      if (!equilibrium_reached) message("Restarting push perturbation step: equilibrium not reached.")
    }
    
    
    # Store time to equilibrium for this step
    time_to_equilibrium_steps[step] <- step_time
  
  }
  
  if (!equilibrium_reached) {
  message("perturbation 1 - Skipping to the next iteration after ", max_attempts, " attempts.")
  next
}
    
  
  
  
  
  ################ store results
  

  # Calculate mean time to equilibrium across all steps
  mean_time_to_equilibrium <- mean(time_to_equilibrium_steps)

  # Store final results
  df_replicates_pert1$mean_time_to_equilibrium[[sc]] <- mean_time_to_equilibrium
  df_replicates_pert1$traits_eq_push1[[sc]] <- final_traits
  
  
  
  # Evolutionary rate
    df_replicates_pert1$comm_evolr[[sc]] <- colSums(abs(z.change))/nsims
    df_replicates_pert1$v_evolr[[sc]] <- colSums(abs(z.change[,1:n.host, drop=FALSE]))/nsims
    df_replicates_pert1$e_evolr[[sc]]  <- colSums(abs(z.change[,(n.host+1):nsp, drop=FALSE]))/nsims
    
    # mean fluctuation in trait matching
    df_replicates_pert1$v_fcij[[sc]] <-rowSums(cummulative.matching.change)
    df_replicates_pert1$e_fcij[[sc]] <- colSums(cummulative.matching.change)
    df_replicates_pert1$comm_fcij[[sc]] <- c(rowSums(cummulative.matching.change), colSums(cummulative.matching.change))
    
    # fitness load
    
    
    # --- Victims
    simv_load_total <- numeric(length(n.host))
    simv_load_abiot <- numeric(length(n.host))
    simv_load_biot <- numeric(length(n.host))
    
    for (v in 1:n.host) {
  # Filter partners by interaction matrix
  interacting_exploiters <- which(binary.matrix[v, ] == 1)  # Get indices of interacting exploiters
  x_partners <- trait[sim, n.host + interacting_exploiters]  # Traits of interacting exploiters
  
  fitness_load <- fitness_load_victims(
    x_s = trait[sim, v],
    x_0 = teta[v],
    x_partners = x_partners,  # Only interacting exploiters
    epsilon = e  # Pass interaction threshold
  )
  
  simv_load_total[v] <- fitness_load$load
  simv_load_abiot[v] <- fitness_load$abiotic_L
  simv_load_biot[v] <- fitness_load$biotic_L
}
    
    df_replicates_pert1$v_load_total[[sc]] <- simv_load_total
    df_replicates_pert1$v_load_abiot[[sc]] <- simv_load_abiot
    df_replicates_pert1$v_load_biot[[sc]] <- simv_load_biot
    
    # --- Exploiters
    sime_load_total <- numeric(length(n.host))
    sime_load_abiot <- numeric(length(n.host))
    sime_load_biot <- numeric(length(n.host))
    sime_load_network <- numeric(length(n.host))
    index <- 1
    
    for (e in (n.host + 1):nsp) {
      fitness_load <- fitness_load_exploiters(
        x_s = trait[sim, e],
        x_0 = teta[e],
        x_partners = abs(a.matrix[, e])
      )
      
      sime_load_total[index] <- fitness_load$load
      sime_load_abiot[index] <- fitness_load$abiotic_L
      sime_load_biot[index] <- fitness_load$biotic_L
      sime_load_network[index] <- fitness_load$network_L
      index <- index + 1
    }
    
    df_replicates_pert1$e_load_total[[sc]] <- sime_load_total
    df_replicates_pert1$e_load_abiot[[sc]] <- sime_load_abiot
    df_replicates_pert1$e_load_biot[[sc]] <- sime_load_biot
    df_replicates_pert1$e_load_network[[sc]] <- sime_load_network
    
    
    
    
  }



########################
# PULL PERTURBATION
########################



if (simulate_perturbation2 == "yes") {
    # Pulse Perturbation (Displacement in traits)
    trait[1, ] <- trait[1, ] + rnorm(nsp, mean = 0, sd = 2)  # Apply a pulse perturbation
    equilibrium_reached <- FALSE
    # Initialize time-to-equilibrium tracker
    time_to_equilibrium <- NA
    
    attempt <- 0 # attempts for finding an equilibrium
    while (!equilibrium_reached && attempt < max_attempts) {
      attempt <- attempt + 1
      trait <- matrix(0, nrow=nsims, ncol=nsp)   ### store trait values in each timestep for each sp
      z.change <- matrix(0, nrow=nsims, ncol=nsp)  ### store trait changes in each timestep for each sp
      teta <- df_replicates$teta[[sc]] # teta from nonperturbed simulation
      cummulative.matching.change <- matrix(0, nrow=n.host, ncol=n.exploiter)
      stable_timesteps <- 0
      for (sim in 1:(nsims - 1)) {
        a.matrix <- create_a_matrix(trait_vec = trait[sim, ], Fmat = bigmat.F)
        E <- ksi * (teta - trait[sim, ])
        Rh <- calc_Rh_corrected(a_matrix = a.matrix, big_mat = bigmat.F, KSI.b = bio_sel_v, n_host = n.host, e_threshold = e)
        Rp <- calc_Rp_corrected(a_matrix = a.matrix, big_mat = bigmat.F, KSI.b = bio_sel_e, B = b, n_host = n.host, n_exploiter = n.exploiter, nsp = nsp)
        R <- c(Rh, Rp)
        z.change[sim, ] <- her * (R + E)
        trait[sim + 1, ] <- trait[sim, ] + z.change[sim, ]
        
        # Compute Trait evolution
        R <- c(Rh, Rp) # Biotic selection for victims and exploiters
        if( length(R) != length(E)) stop("something is wrong") #test R & E have same length 
        z.change[sim,] <- her * (R + E) 
        trait[sim+1,] <- trait[sim,] + z.change[sim,] 
        
        # Update a.matrix
        dif.traits2 <- outer(trait[sim+1,], trait[sim+1,], FUN= "-") 
        a.matrixsim2 <- bigmat.F*dif.traits2 
        
        # Calculate change in trait matching with time
        matching.change.matrix <- a.matrixsim2[1:n.host, (n.host+1):nsp] -a.matrix[1:n.host, (n.host+1):nsp]
        cummulative.matching.change <- cummulative.matching.change + abs(matching.change.matrix)
        
        # Check for equilibrium
        if (sim >= max_stable_timesteps && isTRUE(all.equal(trait[sim, ], trait[sim - max_stable_timesteps + 1, ], tolerance = 1e-10))) {
          stable_timesteps <- stable_timesteps + 1
          if (stable_timesteps >= max_stable_timesteps) {
            equilibrium_reached <- TRUE
            time_to_equilibrium <- sim
            message("Equilibrium reached after pulse perturbation at timestep ", sim)
            df_replicates_pert2$traits_eq[[sc]] <- trait[sim, ]
            break
          }
        } else {
          stable_timesteps <- 0
        }
      }
      if (!equilibrium_reached) message("Restarting pulse perturbation simulation: equilibrium not reached.")
    }
    
    
    if (!equilibrium_reached) {
  message("perturbation 2 - Skipping to the next iteration after ", max_attempts, " attempts.")
  next
}
    
    
    
    
    ################ store results
    
    
    # save trait values at equilibrium
    df_replicates_pert2$traits_eq_push2[[sc]] <- trait[sim,]
    
    
    # Evolutionary rate
    df_replicates_pert2$comm_evolr[[sc]] <- colSums(abs(z.change))/nsims
    df_replicates_pert2$v_evolr[[sc]] <- colSums(abs(z.change[,1:n.host, drop=FALSE]))/nsims
    df_replicates_pert2$e_evolr[[sc]]  <- colSums(abs(z.change[,(n.host+1):nsp, drop=FALSE]))/nsims
    
    
    # mean fluctuation in trait matching
    df_replicates_pert2$v_fcij[[sc]] <-rowSums(cummulative.matching.change)
    df_replicates_pert2$e_fcij[[sc]] <- colSums(cummulative.matching.change)
    df_replicates_pert2$comm_fcij[[sc]] <- c(rowSums(cummulative.matching.change), colSums(cummulative.matching.change))
    
    # fitness load
    # --- Victims
    simv_load_total <- numeric(length(n.host))
    simv_load_abiot <- numeric(length(n.host))
    simv_load_biot <- numeric(length(n.host))
    
    for (v in 1:n.host) {
  # Filter partners by interaction matrix
  interacting_exploiters <- which(binary.matrix[v, ] == 1)  # Get indices of interacting exploiters
  x_partners <- trait[sim, n.host + interacting_exploiters]  # Traits of interacting exploiters
  
  fitness_load <- fitness_load_victims(
    x_s = trait[sim, v],
    x_0 = teta[v],
    x_partners = x_partners,  # Only interacting exploiters
    epsilon = e  # Pass interaction threshold
  )
  
  simv_load_total[v] <- fitness_load$load
  simv_load_abiot[v] <- fitness_load$abiotic_L
  simv_load_biot[v] <- fitness_load$biotic_L
}
    
    df_replicates_pert2$v_load_total[[sc]] <- simv_load_total
    df_replicates_pert2$v_load_abiot[[sc]] <- simv_load_abiot
    df_replicates_pert2$v_load_biot[[sc]] <- simv_load_biot
    
    # --- Exploiters
    sime_load_total <- numeric(length(n.host))
    sime_load_abiot <- numeric(length(n.host))
    sime_load_biot <- numeric(length(n.host))
    sime_load_network <- numeric(length(n.host))
    index <- 1
    
    for (e in (n.host + 1):nsp) {
      fitness_load <- fitness_load_exploiters(
        x_s = trait[sim, e],
        x_0 = teta[e],
        x_partners = abs(a.matrix[, e])
      )
      
      sime_load_total[index] <- fitness_load$load
      sime_load_abiot[index] <- fitness_load$abiotic_L
      sime_load_biot[index] <- fitness_load$biotic_L
      sime_load_network[index] <- fitness_load$network_L
      index <- index + 1
    }
    
    df_replicates_pert2$e_load_total[[sc]] <- sime_load_total
    df_replicates_pert2$e_load_abiot[[sc]] <- sime_load_abiot
    df_replicates_pert2$e_load_biot[[sc]] <- sime_load_biot
    df_replicates_pert2$e_load_network[[sc]] <- sime_load_network
    
    # time to equilibrium
    df_replicates_pert2$time_to_equilibrium[[sc]] <- time_to_equilibrium
    
    
}


}

df_replicates_backup <- df_replicates
```





#### Add network metrics to the output table

```{r}

df_replicates <- inner_join(df_replicates, df_network_metrics, by = "id_network")
df_replicates_pert1 <- inner_join(df_replicates_pert1, df_network_metrics, by = "id_network")
df_replicates_pert2 <- inner_join(df_replicates_pert2, df_network_metrics, by = "id_network")

# save output

# saveRDS(df_replicates, here::here("output/df_20replicates_22june2025_corrected.rds"))
# saveRDS(df_replicates_pert1, here::here("output/df_20replicates_pert1_3jan_corrected.rds"))
# saveRDS(df_replicates_pert2, here::here("output/df_20replicates_pert2_3jan_corrected.rds"))


# load output

# df_replicates <- readRDS(here::here("output/df_20replicates_3jan.rds"))
# df_replicates_pert1 <- readRDS(here::here("output/df_20replicates_pert1_3jan.rds"))
# df_replicates_pert2 <- readRDS(here::here("output/df_20replicates_pert2_3jan.rds"))

```



# Plots

```{r}

df_replicates <- df_replicates %>%
  mutate(
    
    # evolutionary rate
    mean_comm_evolr = map_dbl(comm_evolr, ~ mean(.x)),
    mean_e_evolr = map_dbl(e_evolr, ~ mean(.x)),
    mean_v_evolr = map_dbl(v_evolr, ~ mean(.x)),
    
    # trait matching fluctuation
    mean_comm_fcij = map_dbl(comm_fcij, ~ mean(.x)),
    mean_e_fcij = map_dbl(e_fcij, ~ mean(.x)),
    mean_v_fcij = map_dbl(v_fcij, ~ mean(.x)),
    
    # fitness load
    mean_comm_load_total = map_dbl(comm_load_total, ~ mean(.x)),
    mean_comm_load_abiot = map_dbl(comm_load_abiot, ~ mean(.x)),
    mean_comm_load_biot = map_dbl(comm_load_biot, ~ mean(.x)),
    mean_comm_load_network = map_dbl(comm_load_network, ~ mean(.x)),
    
    mean_v_load_total = map_dbl(v_load_total, ~ mean(.x)),
    mean_v_load_abiot = map_dbl(v_load_abiot, ~ mean(.x)),
    mean_v_load_biot = map_dbl(v_load_biot, ~ mean(.x)),
    mean_v_load_network = map_dbl(v_load_network, ~ mean(.x)),
    
    mean_e_load_total = map_dbl(e_load_total, ~ mean(.x)),
    mean_e_load_abiot = map_dbl(e_load_abiot, ~ mean(.x)),
    mean_e_load_biot = map_dbl(e_load_biot, ~ mean(.x)),
    mean_e_load_network = map_dbl(e_load_network, ~ mean(.x))
  )



# df_replicates_pert1 <- df_replicates_pert1 %>%
#   mutate(
#     
#     # evolutionary rate
#     mean_comm_evolr = map_dbl(comm_evolr, ~ mean(.x)),
#     mean_e_evolr = map_dbl(e_evolr, ~ mean(.x)),
#     mean_v_evolr = map_dbl(v_evolr, ~ mean(.x)),
#     
#     # trait matching fluctuation
#     mean_comm_fcij = map_dbl(comm_fcij, ~ mean(.x)),
#     mean_e_fcij = map_dbl(e_fcij, ~ mean(.x)),
#     mean_v_fcij = map_dbl(v_fcij, ~ mean(.x)),
#     
#     # fitness load
#     mean_comm_load_total = map_dbl(comm_load_total, ~ mean(.x)),
#     mean_comm_load_abiot = map_dbl(comm_load_abiot, ~ mean(.x)),
#     mean_comm_load_biot = map_dbl(comm_load_biot, ~ mean(.x)),
#     mean_comm_load_network = map_dbl(comm_load_network, ~ mean(.x)),
#     
#     mean_v_load_total = map_dbl(v_load_total, ~ mean(.x)),
#     mean_v_load_abiot = map_dbl(v_load_abiot, ~ mean(.x)),
#     mean_v_load_biot = map_dbl(v_load_biot, ~ mean(.x)),
#     mean_v_load_network = map_dbl(v_load_network, ~ mean(.x)),
#     
#     mean_e_load_total = map_dbl(e_load_total, ~ mean(.x)),
#     mean_e_load_abiot = map_dbl(e_load_abiot, ~ mean(.x)),
#     mean_e_load_biot = map_dbl(e_load_biot, ~ mean(.x)),
#     mean_e_load_network = map_dbl(e_load_network, ~ mean(.x))
#   )
# 
# 
# df_replicates_pert2 <- df_replicates_pert2 %>%
#   mutate(
#     
#     # evolutionary rate
#     mean_comm_evolr = map_dbl(comm_evolr, ~ mean(.x)),
#     mean_e_evolr = map_dbl(e_evolr, ~ mean(.x)),
#     mean_v_evolr = map_dbl(v_evolr, ~ mean(.x)),
#     
#     # trait matching fluctuation
#     mean_comm_fcij = map_dbl(comm_fcij, ~ mean(.x)),
#     mean_e_fcij = map_dbl(e_fcij, ~ mean(.x)),
#     mean_v_fcij = map_dbl(v_fcij, ~ mean(.x)),
#     
#     # fitness load
#     mean_comm_load_total = map_dbl(comm_load_total, ~ mean(.x)),
#     mean_comm_load_abiot = map_dbl(comm_load_abiot, ~ mean(.x)),
#     mean_comm_load_biot = map_dbl(comm_load_biot, ~ mean(.x)),
#     mean_comm_load_network = map_dbl(comm_load_network, ~ mean(.x)),
#     
#     mean_v_load_total = map_dbl(v_load_total, ~ mean(.x)),
#     mean_v_load_abiot = map_dbl(v_load_abiot, ~ mean(.x)),
#     mean_v_load_biot = map_dbl(v_load_biot, ~ mean(.x)),
#     mean_v_load_network = map_dbl(v_load_network, ~ mean(.x)),
#     
#     mean_e_load_total = map_dbl(e_load_total, ~ mean(.x)),
#     mean_e_load_abiot = map_dbl(e_load_abiot, ~ mean(.x)),
#     mean_e_load_biot = map_dbl(e_load_biot, ~ mean(.x)),
#     mean_e_load_network = map_dbl(e_load_network, ~ mean(.x))
#   )


```


## Spp-level measures

```{r}



df_e_spplevel <- data.frame(
                            "betw_e" = unlist(df_replicates$betw_e),
                            "degree_e" = unlist(df_replicates$degree_e),
                            "e_evolr" = unlist(df_replicates$e_evolr),
                            "e_fcij" = unlist(df_replicates$e_fcij),
                            "e_load_total" = unlist(df_replicates$e_load_total),
                            "e_load_biot" = unlist(df_replicates$e_load_biot),
                            "e_load_abiot" = unlist(df_replicates$e_load_abiot),
                            "e_load_network" = unlist(df_replicates$e_load_network))



# df_e_spplevel_pert1 <- data.frame(
#                             "betw_e" = unlist(df_replicates_pert1$betw_e),
#                             "degree_e" = unlist(df_replicates_pert1$degree_e),
#                             "e_evolr" = unlist(df_replicates_pert1$e_evolr),
#                             "e_fcij" = unlist(df_replicates_pert1$e_fcij),
#                             "e_load_total" = unlist(df_replicates_pert1$e_load_total),
#                             "e_load_biot" = unlist(df_replicates_pert1$e_load_biot),
#                             "e_load_abiot" = unlist(df_replicates_pert1$e_load_abiot))


 
# df_e_spplevel_pert2 <- data.frame(
#                             "betw_e" = unlist(df_replicates_pert2$betw_e),
#                             "degree_e" = unlist(df_replicates_pert2$degree_e),
#                             "e_evolr" = unlist(df_replicates_pert2$e_evolr),
#                             "e_fcij" = unlist(df_replicates_pert2$e_fcij),
#                             "e_load_total" = unlist(df_replicates_pert2$e_load_total),
#                             "e_load_biot" = unlist(df_replicates_pert2$e_load_biot),
#                             "e_load_abiot" = unlist(df_replicates_pert2$e_load_abiot),
#                             "e_load_network" = unlist(df_replicates_pert2$e_load_network))





df_v_spplevel <- data.frame(
                            "betw_v" = unlist(df_replicates$betw_v),
                            "degree_v" = unlist(df_replicates$degree_v),
                            "v_evolr" = unlist(df_replicates$v_evolr),
                            "v_fcij" = unlist(df_replicates$v_fcij),
                            "v_load_total" = unlist(df_replicates$v_load_total),
                            "v_load_biot" = unlist(df_replicates$v_load_biot),
                            "v_load_abiot" = unlist(df_replicates$v_load_abiot))


# df_v_spplevel_pert1 <- data.frame(
#                             "betw_v" = unlist(df_replicates_pert1$betw_v),
#                             "degree_v" = unlist(df_replicates_pert1$degree_v),
#                             "v_evolr" = unlist(df_replicates_pert1$v_evolr),
#                             "v_fcij" = unlist(df_replicates_pert1$v_fcij),
#                             "v_load_total" = unlist(df_replicates_pert1$v_load_total),
#                             "v_load_biot" = unlist(df_replicates_pert1$v_load_biot),
#                             "v_load_abiot" = unlist(df_replicates_pert1$v_load_abiot))
# 


# df_v_spplevel_pert2 <- data.frame(
#                             "betw_v" = unlist(df_replicates_pert2$betw_v),
#                             "degree_v" = unlist(df_replicates_pert2$degree_v),
#                             "v_evolr" = unlist(df_replicates_pert2$v_evolr),
#                             "v_fcij" = unlist(df_replicates_pert2$v_fcij),
#                             "v_load_total" = unlist(df_replicates_pert2$v_load_total),
#                             "v_load_biot" = unlist(df_replicates_pert2$v_load_biot),
#                             "v_load_abiot" = unlist(df_replicates_pert2$v_load_abiot))

```



# PLOTS

Pars for plotting

```{r}

my_theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5))


back_transform_sqrt <- function(x) {
  return(x^2)  # Square back the transformed value
}


back_transform_log <- function(x) {
  return(exp(x))  # Exponentiate the log-transformed value
}

format_scientific <- function(x) {
  ifelse(abs(x) < 0.001 | abs(x) > 10000,  # Use | for vectorized logical operation
         sprintf("%.1e", x),  # Scientific notation with one decimal
         round(x, 1))         # Otherwise, round to one decimal
}



```




# Fig2


## exploiters

```{r}

# Evolutonary rate


df_plot <- df_replicates %>% 
  select(id_network, replicate, complexity_index, modularity, NODF, mean_e_evolr, mean_e_fcij) %>%
  group_by(id_network) %>%
  summarize(mean_mean_e_evolr = mean(mean_e_evolr),
            mean_mean_e_fcij = mean(mean_e_fcij),
            complexity_index = complexity_index[1],
            modularity = modularity[1],
            NODF = NODF[1])

# Compute breaks
breaks <- quantile(df_plot$mean_mean_e_evolr, probs = c(0, 1), na.rm = TRUE)

# Compute labels
labels <- c("6.6e-6", "2.7e-2")



plot_complIndex_evolrate <- ggplot(data = df_plot, aes(x = complexity_index, y = mean_mean_e_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
  labs(x = expression(sqrt(C * S)), y = expression(paste("Evolutionary rate ",italic((sigma))))) +
  theme_classic()+
  my_theme

plot_modnest_evolrate <- ggplot(data = df_plot, aes(x = modularity, y = NODF)) +
  geom_point(aes(color = mean_mean_e_evolr), alpha = 0.5, size = 4) +
  labs(x = "Modularity", y = expression(paste("Nestedness ", italic(NODF)))) +
  scale_color_viridis_c(
    name = expression(italic(sigma)),
    option = "turbo"
  ) +
  coord_cartesian(clip = "off") + 
  theme_classic() +
  theme(
    legend.position = c(0.90, 0.85),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = "white", color = "gray70"),
    legend.key.size = unit(0.5, "cm"),
    legend.title.align = 0.5
  ) +
  my_theme


plot_degree_evolrate <- ggplot(data = df_e_spplevel, aes(x = degree_e, y = e_evolr)) +
  geom_point(alpha = 0.03, color = "#E69F00") +
  labs(x = "Species degree", y = expression(paste("Evolutionary rate ",italic((sigma))))) +
  theme_classic()+
  my_theme


# Fluctuation in trait matching

plot_complIndex_fc <- ggplot(data = df_plot, aes(x = complexity_index, y = sqrt(mean_mean_e_fcij))) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
    scale_y_continuous(labels = back_transform_sqrt)+
  labs(x = expression(sqrt(C * S)), y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

# Define breaks and labels for the color scale
breaks <- quantile(sqrt(df_plot$mean_mean_e_fcij), probs = c(0, 1))
labels <- c(round(min(df_plot$mean_mean_e_fcij),1), round(max(df_plot$mean_mean_e_fcij), 1))

plot_modnest_fc <- ggplot(data = df_plot, aes(x = modularity, y = NODF)) +
  geom_point(aes(color = sqrt(mean_mean_e_fcij)), alpha = 0.5, size = 4) +
  labs(x = "Modularity", y = expression(paste("Nestedness ", italic(NODF)))) +
  scale_color_viridis_c(
    name = expression(italic(lambda)),
    option = "turbo"
  ) +
  coord_cartesian(clip = "off") + 
  theme_classic() +
  theme(
    legend.position = c(0.90, 0.85),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = "white", color = "gray70"),
    legend.key.size = unit(0.5, "cm"),
    legend.title.align = 0.5
  ) +
  my_theme



plot_degree_fc <- ggplot(data = df_e_spplevel, aes(x = degree_e, y = sqrt(e_fcij))) +
  geom_point(alpha = 0.03, color = "#E69F00") +
  scale_y_continuous(labels = back_transform_sqrt)+
  labs(x = "Species degree", y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme


```


## victims

```{r}

# Evolutonary rate


df_plot <- df_replicates %>% 
  select(id_network, replicate, complexity_index, modularity, NODF, mean_v_evolr, mean_v_fcij) %>%
  group_by(id_network) %>%
  summarize(mean_mean_v_evolr = mean(mean_v_evolr),
            mean_mean_v_fcij = mean(mean_v_fcij),
            complexity_index = complexity_index[1],
            modularity = modularity[1],
            NODF = NODF[1])

breaks <- quantile(df_plot$mean_mean_v_evolr, probs = c(0, 1))
labels <- c(round(min(df_plot$mean_mean_v_evolr),1), round(max(df_plot$mean_mean_v_evolr), 1))

plot_complIndex_evolrate_v <- ggplot(data = df_plot, aes(x = complexity_index, y = mean_mean_v_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
  labs(x = expression(sqrt(C * S)), y = expression(paste("Evolutionary rate ",italic((sigma))))) +
  theme_classic()+
  my_theme

plot_modnest_evolrate_v <- ggplot(data = df_plot, aes(x = modularity, y = NODF)) +
  geom_point(aes(color = mean_mean_v_evolr), alpha = 0.5, size = 4) +
  labs(x = "Modularity", y = expression(paste("Nestedness ", italic(NODF)))) +
  scale_color_viridis_c(
    name = expression(italic(sigma)),
    option = "turbo"
  ) +
  coord_cartesian(clip = "off") +  
  theme_classic() +
  theme(
    legend.position = c(0.90, 0.85),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = "white", color = "gray70"),
    legend.key.size = unit(0.5, "cm"),
    legend.title.align = 0.5
  ) +
  my_theme



plot_degree_evolrate_v <- ggplot(data = df_v_spplevel, aes(x = degree_v, y = v_evolr)) +
  geom_point(alpha = 0.03, color = "#009E73") +
  labs(x = "Species degree", y = expression(paste("Evolutionary rate ",italic((sigma))))) +
  theme_classic()+
  my_theme


# Fluctuation in trait matching

plot_complIndex_fc_v <- ggplot(data = df_plot, aes(x = complexity_index, y = sqrt(mean_mean_v_fcij))) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
  scale_y_continuous(labels = back_transform_sqrt)+
  labs(x = expression(sqrt(C * S)), y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

# Define breaks and labels for the color scale

breaks <- quantile(sqrt(df_plot$mean_mean_v_fcij), probs = c(0, 1))
labels <- c(round(min(df_plot$mean_mean_v_fcij),1), round(max(df_plot$mean_mean_v_fcij), 1))

plot_modnest_fc_v <- ggplot(data = df_plot, aes(x = modularity, y = NODF)) +
  geom_point(aes(color = sqrt(mean_mean_v_fcij)), alpha = 0.5, size = 4) +
  labs(x = "Modularity", y = expression(paste("Nestedness ", italic(NODF)))) +
  scale_color_viridis_c(
    name = expression(italic(lambda)),
    labels = back_transform_sqrt,
    option = "turbo"
  ) +
  coord_cartesian(clip = "off") + 
  theme_classic() +
  theme(
    legend.position = c(0.90, 0.85),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = "white", color = "gray70"),
    legend.key.size = unit(0.5, "cm"),
    legend.title.align = 0.5
  ) +
  my_theme


plot_degree_fc_v <- ggplot(data = df_v_spplevel, aes(x = degree_v, y = sqrt(v_fcij))) +
  geom_point(alpha = 0.03, color = "#009E73") +
  scale_y_continuous(labels = back_transform_sqrt)+
  labs(x = "Species degree", y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme


```



# Fig3


```{r}
fig_3_rate_e <- ggarrange(
  plot_complIndex_evolrate, plot_modnest_evolrate, plot_degree_evolrate,
  nrow = 1,
  ncol = 3,
  labels = LETTERS[1:3]
)

fig_3_rate_v <- ggarrange(
  plot_complIndex_evolrate_v, plot_modnest_evolrate_v, plot_degree_evolrate_v,
  nrow = 1,
  ncol = 3,
  labels = LETTERS[4:6]
)


fig_3_cf_e <- ggarrange(
  plot_complIndex_fc, plot_modnest_fc, plot_degree_fc,
  nrow = 1,
  ncol = 3,
  labels = LETTERS[7:9]
)

fig_3_cf_v <- ggarrange(
  plot_complIndex_fc_v, plot_modnest_fc_v, plot_degree_fc_v,
  nrow = 1,
  ncol = 3,
  labels = LETTERS[10:12]
)



fig_3 <- ggarrange(
  
  fig_3_rate_e,
  fig_3_rate_v,
  fig_3_cf_e,
  fig_3_cf_v,
  
  nrow = 4,
  ncol = 1
)

fig_3


#ggsave(here::here("figures/fig_3_15Aug2025.png"), height = 16, width = 12)

```





# Fig4


```{r}


plot_adaptation_load_e <-  ggplot(data = df_e_spplevel, aes(x = e_load_abiot, y = sqrt(e_load_biot))) +  
  geom_point(aes(color = sqrt(e_load_network)), alpha = 0.5, size = 3) + 
  labs(x = expression(paste("Abiotic load ", italic((L[alpha])))), y = expression(paste("Biotic load ", italic((L[beta]))))) +
  scale_y_continuous(labels = back_transform_sqrt)+
  scale_color_viridis_c(name = expression(paste("Network load ", italic((L[gamma])))), labels = back_transform_sqrt, option = "turbo") +
  theme_classic()+
  theme(legend.position = "top",
        legend.key.size = unit(0.5, "cm"))+
  my_theme

plot_adaptation_load_v <-  ggplot(data = df_v_spplevel, aes(x = v_load_abiot, y = sqrt(v_load_biot))) +  
  geom_point(alpha = 0.5, size = 3) + 
  labs(x = expression(paste("Abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
  scale_y_continuous(labels = back_transform_sqrt)+
  scale_color_viridis_c(name = expression(paste("Network load ", italic((L[gamma])))), labels = back_transform_sqrt, option = "turbo") +
  theme_classic()+
  theme(legend.position = "top",
        legend.key.size = unit(0.5, "cm"))+
  my_theme

plot_load_components = ggarrange(
  
  plot_adaptation_load_e,
  plot_adaptation_load_v,
  
  nrow = 2,
  ncol = 1,
  
  common.legend = TRUE,
  
  legend = "right",
  
  labels = LETTERS[1:2]
  
)

plot_load_components

```


```{r}
df_plot_e <- df_replicates %>% 
  select(id_network, replicate, complexity_index, modularity, NODF, mean_e_load_abiot, mean_e_load_biot, mean_e_load_network) %>%
  group_by(id_network) %>%
  summarize(mean_mean_e_load_abiot = mean(mean_e_load_abiot),
            mean_mean_e_load_biot = mean(mean_e_load_biot),
            mean_mean_e_load_network = mean(mean_e_load_network),
            complexity_index = complexity_index[1],
            modularity = modularity[1],
            NODF = NODF[1])

df_plot_v <- df_replicates %>% 
  select(id_network, replicate, complexity_index, modularity, NODF, mean_v_load_abiot, mean_v_load_biot, mean_v_load_network) %>%
  group_by(id_network) %>%
  summarize(mean_mean_v_load_abiot = mean(mean_v_load_abiot),
            mean_mean_v_load_biot = mean(mean_v_load_biot),
            mean_mean_v_load_network = mean(mean_v_load_network),
            complexity_index = complexity_index[1],
            modularity = modularity[1],
            NODF = NODF[1])
```



```{r}
plot_biotic_complexity_exploiters <- ggplot(data = df_plot_e, aes(x = complexity_index, y = sqrt(mean_mean_e_load_biot))) +
  geom_point(alpha = 0.5, size = 3, color = "#E69F00") +
  scale_y_continuous(labels = back_transform_sqrt) +
  labs(x = expression(sqrt(C * S)), y = expression(paste("Biotic load (", italic(L^{biotic}), ")"))) +
  theme_classic() +
  my_theme

plot_network_complexity_exploiters <- ggplot(data = df_plot_e, aes(x = complexity_index, y = sqrt(mean_mean_e_load_network))) +
  geom_point(alpha = 0.5, size = 3, color = "#E69F00") +
  scale_y_continuous(labels = back_transform_sqrt) +
  labs(x = expression(sqrt(C * S)), y = expression(paste("Network load (", italic(L^{network}), ")"))) +
  theme_classic() +
  my_theme

plot_biotic_complexity_victims <- ggplot(data = df_plot_v, aes(x = complexity_index, y = sqrt(mean_mean_v_load_biot))) +
  geom_point(alpha = 0.5, size = 3, color = "#009E73") +
  scale_y_continuous(labels = back_transform_sqrt) +
  labs(x = expression(sqrt(C * S)), y = expression(paste("Biotic load (", italic(L^{biotic}), ")"))) +
  theme_classic() +
  my_theme

plot_biotic_centrality_exploiters <- ggplot(data = df_e_spplevel, aes(x = sqrt(degree_e), y = sqrt(e_load_biot))) +
  geom_point(alpha = 0.1, size = 1, color = "#E69F00") +
  scale_y_continuous(labels = back_transform_sqrt) +
  scale_x_continuous(labels = back_transform_sqrt) +
  labs(x = "Species degree", y = expression(paste("Biotic load (", italic(L^{biotic}), ")"))) +
  theme_classic() +
  my_theme

plot_network_centrality_exploiters <- ggplot(data = df_e_spplevel, aes(x = sqrt(degree_e), y = sqrt(e_load_network))) +
  geom_point(alpha = 0.1, size = 1, color = "#E69F00") +
  scale_y_continuous(labels = back_transform_sqrt) +
  scale_x_continuous(labels = back_transform_sqrt) +
  labs(x = "Species degree", y = expression(paste("Network load (", italic(L^{network}), ")"))) +
  theme_classic() +
  my_theme

plot_biotic_centrality_victims <- ggplot(data = df_v_spplevel, aes(x = sqrt(degree_v), y = sqrt(v_load_biot))) +
  geom_point(alpha = 0.1, size = 1, color = "#009E73") +
  scale_y_continuous(labels = back_transform_sqrt) +
  scale_x_continuous(labels = back_transform_sqrt) +
  labs(x = "Species degree", y = expression(paste("Biotic load (", italic(L^{biotic}), ")"))) +
  theme_classic() +
  my_theme




```




```{r}
fig4 <- ggarrange(
  plot_biotic_complexity_exploiters,
  plot_network_complexity_exploiters,
  plot_biotic_complexity_victims,
  plot_biotic_centrality_exploiters,
  plot_network_centrality_exploiters,
  plot_biotic_centrality_victims,
  ncol = 3,
  nrow = 2,
  labels = LETTERS[1:6]
)

fig4

#ggsave("../figures/fig_4_15Aug2025.png", height = 9, width = 12)
```
