---
title: "R Notebook"
---

This script runs coevolutionary simulations on each simulated network and generates figures illustrating the relationships between network structural properties and measures of coevolution.

The simulation includes an option to introduce abiotic perturbations by shifting species' environmental optima. These perturbation scenarios were not used in the article **Trapped in the web: network architectures spread coevolution and shape adaptation**. The simulation function includes an argument that controls whether such perturbations are applied and is currently set to exclude them.




```{r}
library(tidyverse)
library(ggplot2)
library(reshape2)
library(tidyr)
library(dplyr)
library(ggpubr)
library(viridis)
library(grid)
library(bipartite) # to use the networklevel() function to calculate modularity and nestedness
library(igraph)

# Custom functions

source(here::here("functions/Functions.R"))
source(here::here("functions/load_computation.R"))

```




## prepare df scenarios

Read empirical networks table

```{r}

df_network_metrics <- readRDS(here::here("output/df_metrics_simulated_nets.rds"))

df_replicates <- df_network_metrics %>%
 slice(rep(row_number(), each = 15))

```




## set storing objects

```{r}


df_replicates <- df_replicates %>% 
  mutate(
  
  
# trait values at equilibrium
    
traits_eq = vector("list", nrow(df_replicates)),
teta = vector("list", nrow(df_replicates)),



# evolutionary rate

comm_evolr = vector("list", nrow(df_replicates)),
v_evolr = vector("list", nrow(df_replicates)),
e_evolr = vector("list", nrow(df_replicates)),


# mean fluctuation in trait matching

comm_fcij = vector("list", nrow(df_replicates)),
v_fcij = vector("list", nrow(df_replicates)),
e_fcij = vector("list", nrow(df_replicates)),
    

# mean fitness load

comm_load_total = vector("list", nrow(df_replicates)),
comm_load_biot = vector("list", nrow(df_replicates)),
comm_load_abiot = vector("list", nrow(df_replicates)),
comm_load_network = vector("list", nrow(df_replicates)),

v_load_total = vector("list", nrow(df_replicates)),
v_load_biot = vector("list", nrow(df_replicates)),
v_load_abiot = vector("list", nrow(df_replicates)),
v_load_network = vector("list", nrow(df_replicates)),

e_load_total = vector("list", nrow(df_replicates)),
e_load_biot = vector("list", nrow(df_replicates)),
e_load_abiot = vector("list", nrow(df_replicates)),
e_load_network = vector("list", nrow(df_replicates)),


)


```



## create dataframes for perturbation scenarios

```{r}
#df_replicates_pert1 = df_replicates 
#df_replicates_pert2 = df_replicates 
```


## set fixed parameters

```{r}

nsims <- 300
her <- 0.22 # heritability
e <- 0.5 # # e (epsilon trait distance)
b <- 10 # b (b expoent exploiter preference)


# Set values of biotic selection

bio_sel_v<-0.16
bio_sel_e<-0.16

ksi_hosts<-1-bio_sel_v
ksi_exploiters<- 1-bio_sel_e


# simulate perturbations?

simulate_perturbation1 = "no"
simulate_perturbation2 = "no"

```


## simulation


```{r}


set.seed(1000)


for (sc in 1:nrow(df_replicates)) {
  
  
  print(paste(sprintf("Right now I am doing iteration %s", sc), "of", nrow(df_replicates)))
  
  
  
  
  # get the network
  
  binary.matrix <- unname(as.matrix(df_replicates$binmat[sc][[1]]))
  
  # add at least one interaction to those species disconnected
  
  binary.matrix <- assign_interactions(binary.matrix)
  
  
  bigmat.F <- create_F(binary.matrix)
  
  # set the number of species
  
  n.host <- nrow(binary.matrix)
  n.exploiter <- ncol(binary.matrix)
  nsp <- n.host + n.exploiter
  
  
  # store objects for simulation
  
  trait <- matrix(0, nrow=nsims, ncol=nsp)   ### store trait values in each timestep for each sp
  z.change <- matrix(0, nrow=nsims, ncol=nsp)  ### store trait changes in each timestep for each sp
  trait[1,] <- rnorm(nsp, mean=0, sd=0.1) #Initial trait value randomly sampled
  teta <- trait[1,] #environmental obtimum = initial z value
  df_replicates$teta[[sc]] <- teta # save teta
  cummulative.matching.change <- matrix(0, nrow=n.host, ncol=n.exploiter)
  ksi<-c(rep(ksi_hosts, n.host),rep(ksi_exploiters, n.exploiter)) # set strengh of abiotic selection vector
  
  
  
 
   #######################################################
  
# Define the maximum number of timesteps without change to consider equilibrium
max_stable_timesteps <- 20

# Initialize a counter to keep track of consecutive timesteps with no change
stable_timesteps <- 0

# Initialize a matrix to store the trait values of the previous timestep
previous_traits <- matrix(NA, nrow = nsims, ncol = nsp)

# Copy the trait values from the first timestep to the previous_traits matrix
previous_traits[1, ] <- trait[1, ]
  
  
  
  for (sim in 1:(nsims-1)) {
    
    
        # Store the trait values before updating
    previous_traits[sim + 1, ] <- trait[sim, ]
    
    
    a.matrix <- create_a_matrix(trait_vec = trait[sim,], Fmat = bigmat.F)  # Create matrix of differences in trait values for those species that interact
    
    E <- ksi * (teta - trait[sim,]) #Calculate env.sel for each sp
    
    
    Rh<-calc_Rh(a_matrix = a.matrix, big_mat = bigmat.F,KSI.b = bio_sel_v, n_host = n.host, e_threshold = e)  # Calculate interaction selection for victims & store it in vector Rh
    Rp<-calc_Rp(a_matrix= a.matrix, big_mat=bigmat.F, KSI.b = bio_sel_e, B=b, n_host=n.host, n_exploiter=n.exploiter,nsp=nsp)  # Calculate interaction selection for exploiters & store it in vector Rp
    
    
  
    # Compute Trait evolution
    
    R <- c(Rh, Rp) # Biotic selection for victims and exploiters
    if( length(R) != length(E)) stop("something is wrong") #test R & E have same length 
    z.change[sim,] <- her * (R + E) 
    trait[sim+1,] <- trait[sim,] + z.change[sim,] 
    
    
    
    # Update a.matrix
    
    dif.traits2 <- outer(trait[sim+1,], trait[sim+1,], FUN= "-") 
    a.matrixsim2 <- bigmat.F*dif.traits2 
    
    
    # Calculate change in trait matching with time
    
    matching.change.matrix <- a.matrixsim2[1:n.host, (n.host+1):nsp] -a.matrix[1:n.host, (n.host+1):nsp]
    cummulative.matching.change <- cummulative.matching.change + abs(matching.change.matrix)
    
    

    
    
     # Check if we've reached the stage where we can test for equilibrium
    if (sim >= max_stable_timesteps) {
        # Check if all trait values are the same as the previous timestep's trait values
        if (isTRUE(all.equal(trait[sim, ], previous_traits[sim, ], tolerance = 1e-10))) {
            stable_timesteps <- stable_timesteps + 1
           # cat("Stable timesteps:", stable_timesteps, "\n")
        } else {
            # Reset the stable_timesteps counter if there's a change
            stable_timesteps <- 0
        }
        
        # If trait values have remained unchanged for the specified number of timesteps, stop the simulation
        if (stable_timesteps >= max_stable_timesteps) {
            message("Equilibrium reached. Stopping simulation at timestep ", sim)
            break
        }
    }
    
    
  }
  
  
  #######################################################
  
  
  # save trait values at equilibrium
  
df_replicates$traits_eq[[sc]] <- trait[nrow(trait),]
  
    
  # Evolutionary rate
  
  
  df_replicates$comm_evolr[[sc]] <- colSums(abs(z.change))/nsims
  df_replicates$v_evolr[[sc]] <- colSums(abs(z.change[,1:n.host]))/nsims
  df_replicates$e_evolr[[sc]]  <- colSums(abs(z.change[,(n.host+1):nsp]))/nsims
  
  
  
    
  # mean fluctuation in trait matching
  
  df_replicates$v_fcij[[sc]] <-rowSums(cummulative.matching.change)
  df_replicates$e_fcij[[sc]] <- colSums(cummulative.matching.change)
  df_replicates$comm_fcij[[sc]] <- c(rowSums(cummulative.matching.change), colSums(cummulative.matching.change))
  
  
  
  
  # fitness load
  
  ### --- Victims
  
  simv_load_total <- numeric(length(n.host))
  simv_load_abiot <- numeric(length(n.host))
  simv_load_biot <- numeric(length(n.host))
  simv_load_network <- numeric(length(n.host))
  
  for (v in 1:n.host) {
    
    simv_load_total[v] <- fitness_load_victims(epsilon = e,
                                               x_s = trait[nrow(trait), v],
                                               x_0 = teta[v], 
                                               x_partners = abs(a.matrix[,v]))$load
    simv_load_abiot[v] <- fitness_load_victims(epsilon = e, x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$abiotic_L
    simv_load_biot[v] <- fitness_load_victims(epsilon = e, x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$biotic_L

    
  }
  
  
  df_replicates$v_load_total[[sc]] <- simv_load_total
  df_replicates$v_load_abiot[[sc]] <- simv_load_abiot
  df_replicates$v_load_biot[[sc]] <- simv_load_biot
  df_replicates$v_load_network[[sc]] <- simv_load_network
  
  
  ### --- Exploiters
  
  sime_load_total <- numeric(length(n.host))
  sime_load_abiot <- numeric(length(n.host))
  sime_load_biot <- numeric(length(n.host))
  sime_load_network <- numeric(length(n.host))
  
  index <- 1
  
  for (e in (n.host+1):nsp) {
    
    sime_load_total[index] <- fitness_load_exploiters(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$load
    sime_load_abiot[index] <- fitness_load_exploiters(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$abiotic_L
    sime_load_biot[index] <- fitness_load_exploiters(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$biotic_L
    sime_load_network[index] <- fitness_load_exploiters(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$network_L
    
    index = index + 1
    
  }
  
  df_replicates$e_load_total[[sc]] <- sime_load_total
  df_replicates$e_load_abiot[[sc]] <- sime_load_abiot
  df_replicates$e_load_biot[[sc]] <- sime_load_biot
  df_replicates$e_load_network[[sc]] <- sime_load_network
  
  
  ### --- Community
  
  df_replicates$comm_load_total[[sc]] <- c(simv_load_total,sime_load_total)
  df_replicates$comm_load_abiot[[sc]] <- c(simv_load_abiot, sime_load_abiot)
  df_replicates$comm_load_biot[[sc]] <- c(simv_load_biot, sime_load_biot)
  df_replicates$comm_load_network[[sc]] <- c(simv_load_network, sime_load_network)
  
  
  
  
  if(simulate_perturbation1 == "yes"){
  
  
  #################################################################
  #################################################################
  #                       PERTURBATION 1
  #################################################################
  #################################################################
  

  # store objects for simulation
  
  trait <- matrix(0, nrow=nsims, ncol=nsp)   ### store trait values in each timestep for each sp
  z.change <- matrix(0, nrow=nsims, ncol=nsp)  ### store trait changes in each timestep for each sp
  trait[1,] <- df_replicates$traits_eq[[sc]] # traits at equilibrium from nonperturbed simulation
  teta <- df_replicates$teta[[sc]] # teta from nonperturbed simulation
  cummulative.matching.change <- matrix(0, nrow=n.host, ncol=n.exploiter)
  
 
   #######################################################
  
# Define the maximum number of timesteps without change to consider equilibrium
max_stable_timesteps <- 20

# Initialize a counter to keep track of consecutive timesteps with no change
stable_timesteps <- 0

# Initialize a matrix to store the trait values of the previous timestep
previous_traits <- matrix(NA, nrow = nsims, ncol = nsp)

# Copy the trait values from the first timestep to the previous_traits matrix
previous_traits[1, ] <- trait[1, ]
  
  


# PERTURBATION:

teta = sapply(teta, random_change_teta, SD = 2)



  for (sim in 1:(nsims-1)) {

    
    
            # Store the trait values before updating
    previous_traits[sim + 1, ] <- trait[sim, ]
    
    
    a.matrix <- create_a_matrix(trait_vec = trait[sim,], Fmat = bigmat.F)  # Create matrix of differences in trait values for those species that interact
    
    E <- ksi * (teta - trait[sim,]) #Calculate env.sel for each sp
    
    
    Rh<-calc_Rh(a_matrix = a.matrix, big_mat = bigmat.F,KSI.b = bio_sel_v, n_host = n.host, e_threshold = e)  # Calculate interaction selection for victims & store it in vector Rh
    Rp<-calc_Rp(a_matrix= a.matrix, big_mat=bigmat.F, KSI.b = bio_sel_e, B=b, n_host=n.host, n_exploiter=n.exploiter,nsp=nsp)  # Calculate interaction selection for exploiters & store it in vector Rp
    
    
  
    # Compute Trait evolution
    
    R <- c(Rh, Rp) # Biotic selection for victims and exploiters
    if( length(R) != length(E)) stop("something is wrong") #test R & E have same length 
    z.change[sim,] <- her * (R + E) 
    trait[sim+1,] <- trait[sim,] + z.change[sim,] 
    
    
    
    # Update a.matrix
    
    dif.traits2 <- outer(trait[sim+1,], trait[sim+1,], FUN= "-") 
    a.matrixsim2 <- bigmat.F*dif.traits2 
    
    
    # Calculate change in trait matching with time
    
    matching.change.matrix <- a.matrixsim2[1:n.host, (n.host+1):nsp] -a.matrix[1:n.host, (n.host+1):nsp]
    cummulative.matching.change <- cummulative.matching.change + abs(matching.change.matrix)
    
    
    
    
    
     # Check if we've reached the stage where we can test for equilibrium
    if (sim >= max_stable_timesteps) {
        # Check if all trait values are the same as the previous timestep's trait values
        if (isTRUE(all.equal(trait[sim, ], previous_traits[sim, ], tolerance = 1e-10))) {
            stable_timesteps <- stable_timesteps + 1
           # cat("Stable timesteps:", stable_timesteps, "\n")
        } else {
            # Reset the stable_timesteps counter if there's a change
            stable_timesteps <- 0
        }
        
        # If trait values have remained unchanged for the specified number of timesteps, stop the simulation
        if (stable_timesteps >= max_stable_timesteps) {
            message("Equilibrium reached. Stopping simulation at timestep ", sim)
            break
        }
    }
    
    
  }
  
  
  #######################################################
  
  
  # save trait values at equilibrium
  
df_replicates_pert1$traits_eq[[sc]] <- trait[nrow(trait),]
  
    
  # Evolutionary rate
  
  
  df_replicates_pert1$comm_evolr[[sc]] <- colSums(abs(z.change))/nsims
  df_replicates_pert1$v_evolr[[sc]] <- colSums(abs(z.change[,1:n.host, drop=FALSE]))/nsims
  df_replicates_pert1$e_evolr[[sc]]  <- colSums(abs(z.change[,(n.host+1):nsp, drop=FALSE]))/nsims
  
  
  
    
  # mean fluctuation in trait matching
  
  df_replicates_pert1$v_fcij[[sc]] <-rowSums(cummulative.matching.change)
  df_replicates_pert1$e_fcij[[sc]] <- colSums(cummulative.matching.change)
  df_replicates_pert1$comm_fcij[[sc]] <- c(rowSums(cummulative.matching.change), colSums(cummulative.matching.change))
  
  
  
  
  # fitness load
  
  ### --- Victims
  
  simv_load_total <- numeric(length(n.host))
  simv_load_abiot <- numeric(length(n.host))
  simv_load_biot <- numeric(length(n.host))
  simv_load_network <- numeric(length(n.host))
  
  for (v in 1:n.host) {
    
    simv_load_total[v] <- fitness_load(x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$load
    simv_load_abiot[v] <- fitness_load(x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$abiotic_L
    simv_load_biot[v] <- fitness_load(x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$biotic_L
    simv_load_network[v] <- fitness_load(x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$network_L
    
  }
  
  
  df_replicates_pert1$v_load_total[[sc]] <- simv_load_total
  df_replicates_pert1$v_load_abiot[[sc]] <- simv_load_abiot
  df_replicates_pert1$v_load_biot[[sc]] <- simv_load_biot
  df_replicates_pert1$v_load_network[[sc]] <- simv_load_network
  
  
 ### --- Exploiters
  
  sime_load_total <- numeric(length(n.host))
  sime_load_abiot <- numeric(length(n.host))
  sime_load_biot <- numeric(length(n.host))
  sime_load_network <- numeric(length(n.host))
  
  index <- 1
  
  for (e in (n.host+1):nsp) {
    
    sime_load_total[index] <- fitness_load(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$load
    sime_load_abiot[index] <- fitness_load(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$abiotic_L
    sime_load_biot[index] <- fitness_load(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$biotic_L
    sime_load_network[index] <- fitness_load(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$network_L
    
    index = index + 1
    
  }
  
  df_replicates_pert1$e_load_total[[sc]] <- sime_load_total
  df_replicates_pert1$e_load_abiot[[sc]] <- sime_load_abiot
  df_replicates_pert1$e_load_biot[[sc]] <- sime_load_biot
  df_replicates_pert1$e_load_network[[sc]] <- sime_load_network
  
  ### --- Community
  
  df_replicates_pert1$comm_load_total[[sc]] <- c(simv_load_total,sime_load_total)
  df_replicates_pert1$comm_load_abiot[[sc]] <- c(simv_load_abiot, sime_load_abiot)
  df_replicates_pert1$comm_load_biot[[sc]] <- c(simv_load_biot, sime_load_biot)
  df_replicates_pert1$comm_load_network[[sc]] <- c(simv_load_network, sime_load_network)
  
  
  }
  
  
  
   if(simulate_perturbation2 == "yes"){
  
  
  #################################################################
  #################################################################
  #                       PERTURBATION 2
  #################################################################
  #################################################################
  

  # store objects for simulation
  
  trait <- matrix(0, nrow=nsims, ncol=nsp)   ### store trait values in each timestep for each sp
  z.change <- matrix(0, nrow=nsims, ncol=nsp)  ### store trait changes in each timestep for each sp
  trait[1,] <- df_replicates$traits_eq[[sc]] # traits at equilibrium from nonperturbed simulation
  teta <- df_replicates$teta[[sc]] # teta from nonperturbed simulation
  cummulative.matching.change <- matrix(0, nrow=n.host, ncol=n.exploiter)
  
 
   #######################################################
  
# Define the maximum number of timesteps without change to consider equilibrium
max_stable_timesteps <- 20

# Initialize a counter to keep track of consecutive timesteps with no change
stable_timesteps <- 0

# Initialize a matrix to store the trait values of the previous timestep
previous_traits <- matrix(NA, nrow = nsims, ncol = nsp)

# Copy the trait values from the first timestep to the previous_traits matrix
previous_traits[1, ] <- trait[1, ]
  
  



  for (sim in 1:(nsims-1)) {

    
    # PERTURBATION:
    
    teta[1:n.host] = teta[1:n.host] + 0.01
    
    
            # Store the trait values before updating
    previous_traits[sim + 1, ] <- trait[sim, ]
    
    
    a.matrix <- create_a_matrix(trait_vec = trait[sim,], Fmat = bigmat.F)  # Create matrix of differences in trait values for those species that interact
    
    E <- ksi * (teta - trait[sim,]) #Calculate env.sel for each sp
    
    
    Rh<-calc_Rh(a_matrix = a.matrix, big_mat = bigmat.F,KSI.b = bio_sel_v, n_host = n.host, e_threshold = e)  # Calculate interaction selection for victims & store it in vector Rh
    Rp<-calc_Rp(a_matrix= a.matrix, big_mat=bigmat.F, KSI.b = bio_sel_e, B=b, n_host=n.host, n_exploiter=n.exploiter,nsp=nsp)  # Calculate interaction selection for exploiters & store it in vector Rp
    
    
  
    # Compute Trait evolution
    
    R <- c(Rh, Rp) # Biotic selection for victims and exploiters
    if( length(R) != length(E)) stop("something is wrong") #test R & E have same length 
    z.change[sim,] <- her * (R + E) 
    trait[sim+1,] <- trait[sim,] + z.change[sim,] 
    
    
    
    # Update a.matrix
    
    dif.traits2 <- outer(trait[sim+1,], trait[sim+1,], FUN= "-") 
    a.matrixsim2 <- bigmat.F*dif.traits2 
    
    
    # Calculate change in trait matching with time
    
    matching.change.matrix <- a.matrixsim2[1:n.host, (n.host+1):nsp] -a.matrix[1:n.host, (n.host+1):nsp]
    cummulative.matching.change <- cummulative.matching.change + abs(matching.change.matrix)
    
    
    
    
    
     # Check if we've reached the stage where we can test for equilibrium
    if (sim >= max_stable_timesteps) {
        # Check if all trait values are the same as the previous timestep's trait values
        if (isTRUE(all.equal(trait[sim, ], previous_traits[sim, ], tolerance = 1e-10))) {
            stable_timesteps <- stable_timesteps + 1
            cat("Stable timesteps:", stable_timesteps, "\n")
        } else {
            # Reset the stable_timesteps counter if there's a change
            stable_timesteps <- 0
        }
        
        # If trait values have remained unchanged for the specified number of timesteps, stop the simulation
        if (stable_timesteps >= max_stable_timesteps) {
            message("Equilibrium reached. Stopping simulation at timestep ", sim)
            break
        }
    }
    
    
  }
  
  
  #######################################################
  
  
  # save trait values at equilibrium
  
df_replicates_pert2$traits_eq[[sc]] <- trait[nrow(trait),]
  
    
  # Evolutionary rate
  
  
  df_replicates_pert2$comm_evolr[[sc]] <- colSums(abs(z.change))/nsims
  df_replicates_pert2$v_evolr[[sc]] <- colSums(abs(z.change[,1:n.host, drop=FALSE]))/nsims
  df_replicates_pert2$e_evolr[[sc]]  <- colSums(abs(z.change[,(n.host+1):nsp, drop=FALSE]))/nsims
  
  
  
    
  # mean fluctuation in trait matching
  
  df_replicates_pert2$v_fcij[[sc]] <-rowSums(cummulative.matching.change)
  df_replicates_pert2$e_fcij[[sc]] <- colSums(cummulative.matching.change)
  df_replicates_pert2$comm_fcij[[sc]] <- c(rowSums(cummulative.matching.change), colSums(cummulative.matching.change))
  
  
  
  
  # fitness load
  
  ### --- Victims
  
  simv_load_total <- numeric(length(n.host))
  simv_load_abiot <- numeric(length(n.host))
  simv_load_biot <- numeric(length(n.host))
  simv_load_network <- numeric(length(n.host))
  
  for (v in 1:n.host) {
    
    simv_load_total[v] <- fitness_load(x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$load
    simv_load_abiot[v] <- fitness_load(x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$abiotic_L
    simv_load_biot[v] <- fitness_load(x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$biotic_L
    simv_load_network[v] <- fitness_load(x_s = trait[nrow(trait), v], x_0 = teta[v], x_partners = abs(a.matrix[,v]))$network_L
    
  }
  
  
  df_replicates_pert2$v_load_total[[sc]] <- simv_load_total
  df_replicates_pert2$v_load_abiot[[sc]] <- simv_load_abiot
  df_replicates_pert2$v_load_biot[[sc]] <- simv_load_biot
  df_replicates_pert2$v_load_network[[sc]] <- simv_load_network
  
  
 ### --- Exploiters
  
  sime_load_total <- numeric(length(n.host))
  sime_load_abiot <- numeric(length(n.host))
  sime_load_biot <- numeric(length(n.host))
  sime_load_network <- numeric(length(n.host))
  
  index <- 1
  
  for (e in (n.host+1):nsp) {
    
    sime_load_total[index] <- fitness_load(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$load
    sime_load_abiot[index] <- fitness_load(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$abiotic_L
    sime_load_biot[index] <- fitness_load(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$biotic_L
    sime_load_network[index] <- fitness_load(x_s = trait[nrow(trait), e], x_0 = teta[e], x_partners = abs(a.matrix[,e]))$network_L
    
    index = index + 1
    
  }
  
  df_replicates_pert2$e_load_total[[sc]] <- sime_load_total
  df_replicates_pert2$e_load_abiot[[sc]] <- sime_load_abiot
  df_replicates_pert2$e_load_biot[[sc]] <- sime_load_biot
  df_replicates_pert2$e_load_network[[sc]] <- sime_load_network
  
  ### --- Community
  
  df_replicates_pert2$comm_load_total[[sc]] <- c(simv_load_total,sime_load_total)
  df_replicates_pert2$comm_load_abiot[[sc]] <- c(simv_load_abiot, sime_load_abiot)
  df_replicates_pert2$comm_load_biot[[sc]] <- c(simv_load_biot, sime_load_biot)
  df_replicates_pert2$comm_load_network[[sc]] <- c(simv_load_network, sime_load_network)
  
  
  }
  

  

  
}








```



Save data

```{r}
saveRDS(df_replicates, "../output/df_15replicates_4apr2025dec_simnets.rds")
#saveRDS(df_replicates_pert1, "../output/df_1replicates_pert1_5april_simnets.rds.rds")
#saveRDS(df_replicates_pert2, "../output/df_1replicates_pert2_5april_simnets.rds.rds")


df_replicates <- readRDS("../output/df_15replicates_4apr2025dec_simnets.rds")
#df_replicates_pert1 <- readRDS("../output/df_5replicates_pert1_1april_simnets.rds.rds")
#df_replicates_pert2 <- readRDS("../output/df_5replicates_pert2_1april_simnets.rds.rds")

```



# Plotting

```{r}
df_replicates <- df_replicates %>%
  mutate(
    
    # evolutionary rate
    mean_comm_evolr = map_dbl(comm_evolr, ~ mean(.x, na.rm = TRUE)),
    mean_e_evolr = map_dbl(e_evolr, ~ mean(.x, na.rm = TRUE)),
    mean_v_evolr = map_dbl(v_evolr, ~ mean(.x, na.rm = TRUE)),
    
    # trait matching fluctuation
    mean_comm_fcij = map_dbl(comm_fcij, ~ mean(.x, na.rm = TRUE)),
    mean_e_fcij = map_dbl(e_fcij, ~ mean(.x, na.rm = TRUE)),
    mean_v_fcij = map_dbl(v_fcij, ~ mean(.x, na.rm = TRUE)),
    
    # fitness load
    mean_comm_load_total = map_dbl(comm_load_total, ~ mean(.x, na.rm = TRUE)),
    mean_comm_load_abiot = map_dbl(comm_load_abiot, ~ mean(.x, na.rm = TRUE)),
    mean_comm_load_biot = map_dbl(comm_load_biot, ~ mean(.x, na.rm = TRUE)),
    mean_comm_load_network = map_dbl(comm_load_network, ~ mean(.x, na.rm = TRUE)),
    
    mean_v_load_total = map_dbl(v_load_total, ~ mean(.x, na.rm = TRUE)),
    mean_v_load_abiot = map_dbl(v_load_abiot, ~ mean(.x, na.rm = TRUE)),
    mean_v_load_biot = map_dbl(v_load_biot, ~ mean(.x, na.rm = TRUE)),
    mean_v_load_network = map_dbl(v_load_network, ~ mean(.x, na.rm = TRUE)),
    
    mean_e_load_total = map_dbl(e_load_total, ~ mean(.x, na.rm = TRUE)),
    mean_e_load_abiot = map_dbl(e_load_abiot, ~ mean(.x, na.rm = TRUE)),
    mean_e_load_biot = map_dbl(e_load_biot, ~ mean(.x, na.rm = TRUE)),
    mean_e_load_network = map_dbl(e_load_network, ~ mean(.x, na.rm = TRUE))
  )


```



Pars for plotting



```{r}

my_theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5))


back_transform_sqrt <- function(x) {
  return(x^2)  # Square back the transformed value
}


# Define the desired order of levels
custom_order <- c("random", "scale-free", "nested")  # Replace with your desired order

# Convert distribution to a factor with custom order
df_replicates$distribution <- factor(df_replicates$distribution, levels = custom_order)

```




# evol rate & matching ~ structure


## richness

```{r}
df_plot <- df_replicates %>% 
  filter(test_metric == "richness") %>% 
  select(connectance, n,
         mean_e_evolr, mean_v_evolr, mean_comm_evolr,
         mean_e_fcij, mean_v_fcij, mean_comm_fcij,
         NODF,
         modularity) %>%
  group_by(n) %>%
  summarize(mean_mean_e_evolr = mean(mean_e_evolr),
            mean_mean_v_evolr = mean(mean_v_evolr),
            mean_mean_comm_evolr = mean(mean_comm_evolr),
            mean_mean_e_fcij = mean(mean_e_fcij),
            mean_mean_v_fcij = mean(mean_v_fcij),
            mean_mean_comm_fcij = mean(mean_comm_fcij),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])


s_er_e = ggplot(data = df_plot, aes(x = n, y = mean_mean_e_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species richness", y = expression(paste("Evolutionary rate ",italic((sigma))^2))) +
  theme_classic()+
  my_theme

s_er_v = ggplot(data = df_plot, aes(x = n, y = mean_mean_v_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species richness", y = expression(paste("Evolutionary rate ",italic((sigma))^2))) +
  theme_classic()+
  my_theme

s_s_e = ggplot(data = df_plot, aes(x = n, y = mean_mean_e_fcij)) +
  geom_point(alpha = 0.3, size = 3,color = "#E69F00") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species richness", y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

s_s_v = ggplot(data = df_plot, aes(x = n, y = mean_mean_v_fcij)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species richness", y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

df_combined <- df_plot %>%
  select(n, mean_mean_e_fcij, mean_mean_v_fcij) %>%
  pivot_longer(
    cols = c(mean_mean_e_fcij, mean_mean_v_fcij),
    names_to = "group",
    values_to = "value"
  ) %>%
  mutate(group = recode(group, 
                        mean_mean_e_fcij = "Exploiters", 
                        mean_mean_v_fcij = "Victims"))

s_s_combined <- ggplot(df_combined, aes(x = n, y = value, color = group)) +
  geom_point(alpha = 0.6, size = 3) +
  labs(
    x = "Species richness",
    y = expression(paste("Trait matching fluctuation ", (italic(lambda)))),
    color = NULL
  ) +
  scale_color_manual(values = c("Exploiters" = "#E69F00", "Victims" = "#009E73")) +
  theme_classic() +
  my_theme


s_er_matching = ggarrange(
  
  s_er_e, s_er_v,
  s_s_e, s_s_v,
  
  labels = LETTERS[1:4],
  
  ncol = 4,
  nrow = 1
)

```




## connectance

```{r}
df_plot <- df_replicates %>% 
  filter(test_metric == "connectance") %>% 
  select(connectance, 
         mean_e_evolr, mean_v_evolr, mean_comm_evolr,
         mean_e_fcij, mean_v_fcij, mean_comm_fcij,
         NODF,
         modularity) %>%
  group_by(connectance) %>%
  summarize(mean_mean_e_evolr = mean(mean_e_evolr),
            mean_mean_v_evolr = mean(mean_v_evolr),
            mean_mean_comm_evolr = mean(mean_comm_evolr),
            mean_mean_e_fcij = mean(mean_e_fcij),
            mean_mean_v_fcij = mean(mean_v_fcij),
            mean_mean_comm_fcij = mean(mean_comm_fcij),
            modularity = modularity[1],
            NODF = NODF[1])


c_er_e = ggplot(data = df_plot, aes(x = connectance, y = mean_mean_e_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Connectance", y = expression(paste("Evolutionary rate ",italic((sigma))^2))) +
  theme_classic()+
  my_theme

c_er_v = ggplot(data = df_plot, aes(x = connectance, y = mean_mean_v_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Connectance", y = expression(paste("Evolutionary rate ",italic((sigma))^2))) +
  theme_classic()+
  my_theme

c_s_e = ggplot(data = df_plot, aes(x = connectance, y = mean_mean_e_fcij)) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Connectance", y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

c_s_v = ggplot(data = df_plot, aes(x = connectance, y = mean_mean_v_fcij)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Connectance", y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

df_combined <- df_plot %>%
  select(connectance, mean_mean_e_fcij, mean_mean_v_fcij) %>%
  pivot_longer(
    cols = c(mean_mean_e_fcij, mean_mean_v_fcij),
    names_to = "group",
    values_to = "value"
  ) %>%
  mutate(group = recode(group, 
                        mean_mean_e_fcij = "Exploiters", 
                        mean_mean_v_fcij = "Victims"))

c_s_combined <- ggplot(df_combined, aes(x = connectance, y = value, color = group)) +
  geom_point(alpha = 0.6, size = 3) +
  labs(
    x = "Connectance",
    y = expression(paste("Trait matching fluctuation ", (italic(lambda)))),
    color = NULL
  ) +
  scale_color_manual(values = c("Exploiters" = "#E69F00", "Victims" = "#009E73")) +
  theme_classic() +
  my_theme


c_er_matching = ggarrange(
  
  c_er_e, c_er_v,
  c_s_e, c_s_v,
  
  labels = LETTERS[5:8],
  
  ncol = 4,
  nrow = 1
)

```


## nestedness

```{r}
df_plot <- df_replicates %>% 
  filter(test_metric == "nestedness") %>% 
  select(connectance, nestedness,
         mean_e_evolr, mean_v_evolr, mean_comm_evolr,
         mean_e_fcij, mean_v_fcij, mean_comm_fcij,
         NODF,
         modularity) %>%
  group_by(nestedness) %>%
  summarize(mean_mean_e_evolr = mean(mean_e_evolr),
            mean_mean_v_evolr = mean(mean_v_evolr),
            mean_mean_comm_evolr = mean(mean_comm_evolr),
            mean_mean_e_fcij = mean(mean_e_fcij),
            mean_mean_v_fcij = mean(mean_v_fcij),
            mean_mean_comm_fcij = mean(mean_comm_fcij),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance = connectance[1])


nest_er_e = ggplot(data = df_plot, aes(x = nestedness, y = mean_mean_e_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("Nestedness ", italic((NODF)))), y = expression(paste("Evolutionary rate ",italic((sigma))^2))) +
  theme_classic()+
  my_theme

nest_er_v = ggplot(data = df_plot, aes(x = nestedness, y = mean_mean_v_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("Nestedness ", italic((NODF)))), y = expression(paste("Evolutionary rate ",italic((sigma))^2))) +
  theme_classic()+
  my_theme

nest_s_e = ggplot(data = df_plot, aes(x = nestedness, y = mean_mean_e_fcij)) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("Nestedness ", italic((NODF)))), y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

nest_s_v = ggplot(data = df_plot, aes(x = nestedness, y = mean_mean_v_fcij)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("Nestedness ", italic((NODF)))), y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

df_combined <- df_plot %>%
  select(nestedness, mean_mean_e_fcij, mean_mean_v_fcij) %>%
  pivot_longer(
    cols = c(mean_mean_e_fcij, mean_mean_v_fcij),
    names_to = "group",
    values_to = "value"
  ) %>%
  mutate(group = recode(group, 
                        mean_mean_e_fcij = "Exploiters", 
                        mean_mean_v_fcij = "Victims"))

nest_s_combined <- ggplot(df_combined, aes(x = nestedness, y = value, color = group)) +
  geom_point(alpha = 0.6, size = 3) +
  labs(
    x = expression(paste("Nestedness ", italic((NODF)))),
    y = expression(paste("Trait matching fluctuation ", (italic(lambda)))),
    color = NULL
  ) +
  scale_color_manual(values = c("Exploiters" = "#E69F00", "Victims" = "#009E73")) +
  theme_classic() +
  my_theme


nest_er_matching = ggarrange(
  
  nest_er_e, nest_er_v,
  nest_s_e, nest_s_v,
  
  labels = LETTERS[9:12],
  
  ncol = 4,
  nrow = 1
)

```




## modularity & nestedness

```{r}
df_plot <- df_replicates %>% 
  filter(test_metric == "nestedness") %>% 
  select(modularity, NODF,
         mean_e_evolr, mean_v_evolr, mean_comm_evolr,
         mean_e_fcij, mean_v_fcij, mean_comm_fcij) %>%
  filter(!is.na(modularity), !is.na(NODF)) |> 
  group_by(modularity) %>%
  summarize(mean_mean_e_evolr = mean(mean_e_evolr),
            mean_mean_v_evolr = mean(mean_v_evolr),
            mean_mean_comm_evolr = mean(mean_comm_evolr),
            mean_mean_e_fcij = mean(mean_e_fcij),
            mean_mean_v_fcij = mean(mean_v_fcij),
            mean_mean_comm_fcij = mean(mean_comm_fcij),
            modularity = modularity[1],
            NODF = NODF[1])


# Plot 2: Evolutionary rate (??˛)
plot_modnest_evolr_e <- ggplot(df_plot, aes(x = modularity, y = NODF, color = mean_mean_e_evolr)) +
  geom_point(size = 4, alpha = 0.8) +
  scale_color_viridis_c(
    name = expression(italic(sigma^2)),
    option = "turbo"
  ) +
  labs(x = "Modularity", y = expression(paste("Nestedness ", italic(NODF)))) +
  theme_classic() +
  theme(
    legend.position = c(0.98, 0.98),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = "white", color = "gray80"),
    legend.key.size = unit(0.5, "cm"),
    legend.title.align = 0.5
  ) +
  my_theme



# Plot 2: Evolutionary rate (??˛)
plot_modnest_evolr_v <- ggplot(df_plot, aes(x = modularity, y = NODF, color = mean_mean_v_evolr)) +
  geom_point(size = 4, alpha = 0.8, shape = 17) +
  scale_color_viridis_c(
    name = expression(italic(sigma^2)),
    option = "turbo"
  ) +
  labs(x = "Modularity", y = expression(paste("Nestedness ", italic(NODF)))) +
  theme_classic() +
  theme(
    legend.position = c(0.98, 0.98),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = "white", color = "gray80"),
    legend.key.size = unit(0.5, "cm"),
    legend.title.align = 0.5
  ) +
  my_theme


# Prepare data for combined plot with both groups
df_combined_modnest <- df_plot %>%
  select(modularity, NODF, mean_mean_e_fcij, mean_mean_v_fcij) %>%
  pivot_longer(
    cols = c(mean_mean_e_fcij, mean_mean_v_fcij),
    names_to = "group",
    values_to = "value"
  ) %>%
  mutate(group = recode(group, 
                        mean_mean_e_fcij = "Exploiters", 
                        mean_mean_v_fcij = "Victims"))

# Create the plot
plot_modnest_fcij_combined <- ggplot(df_combined_modnest, aes(x = modularity, y = NODF, color = value)) +
  geom_point(aes(shape = group), size = 3, alpha = 0.8) +
  scale_color_viridis_c(
    name = expression(italic(lambda)),
    option = "turbo"
  ) +
  scale_shape_manual(values = c("Exploiters" = 16, "Victims" = 17)) +
  guides(shape = "none") +
  labs(x = "Modularity", y = expression(paste("Nestedness ", italic(NODF)))) +
  theme_classic() +
  theme(
    legend.position = c(0.98, 0.98),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = "white", color = "gray80"),
    legend.key.size = unit(0.5, "cm"),
    legend.title.align = 0.5
  ) +
  my_theme



```





## modularity

```{r}
df_plot <- df_replicates %>% 
  filter(test_metric == "nestedness") %>% 
  select(connectance, nestedness,
         mean_e_evolr, mean_v_evolr, mean_comm_evolr,
         mean_e_fcij, mean_v_fcij, mean_comm_fcij,
         NODF,
         modularity) %>%
  group_by(modularity) %>%
  summarize(mean_mean_e_evolr = mean(mean_e_evolr),
            mean_mean_v_evolr = mean(mean_v_evolr),
            mean_mean_comm_evolr = mean(mean_comm_evolr),
            mean_mean_e_fcij = mean(mean_e_fcij),
            mean_mean_v_fcij = mean(mean_v_fcij),
            mean_mean_comm_fcij = mean(mean_comm_fcij),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance = connectance[1])


mod_er_e = ggplot(data = df_plot, aes(x = modularity, y = mean_mean_e_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Modularity", y = expression(paste("Evolutionary rate ",italic((sigma))^2))) +
  theme_classic()+
  my_theme

mod_er_v = ggplot(data = df_plot, aes(x = modularity, y = mean_mean_v_evolr)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Modularity", y = expression(paste("Evolutionary rate ",italic((sigma))^2))) +
  theme_classic()+
  my_theme

mod_s_e = ggplot(data = df_plot, aes(x = modularity, y = mean_mean_e_fcij)) +
  geom_point(alpha = 0.3, size = 3, color = "#E69F00") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Modularity", y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

mod_s_v = ggplot(data = df_plot, aes(x = modularity, y = mean_mean_v_fcij)) +
  geom_point(alpha = 0.3, size = 3, color = "#009E73") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Modularity", y = expression(paste("Trait matching fluctuation ",(italic(lambda))))) +
  theme_classic()+
  my_theme

# Restructure the data for a combined plot
df_combined <- df_plot %>%
  select(modularity, mean_mean_e_fcij, mean_mean_v_fcij) %>%
  pivot_longer(
    cols = c(mean_mean_e_fcij, mean_mean_v_fcij),
    names_to = "group",
    values_to = "value"
  ) %>%
  mutate(group = recode(group, 
                        mean_mean_e_fcij = "Exploiters", 
                        mean_mean_v_fcij = "Victims"))

mod_s_combined <- ggplot(df_combined, aes(x = modularity, y = value, color = group)) +
  geom_point(alpha = 0.6, size = 3) +
  labs(
    x = "Modularity",
    y = expression(paste("Trait matching fluctuation ", (italic(lambda)))),
    color = NULL
  ) +
  scale_color_manual(values = c("Exploiters" = "#E69F00", "Victims" = "#009E73")) +
  theme_classic() +
  my_theme



mod_er_matching = ggarrange(
  
  mod_er_e, mod_er_v,
  mod_s_e, mod_s_v,
  
  labels = LETTERS[9:12],
  
  ncol = 4,
  nrow = 1
)


```

## distribution

```{r}
# Filter out the "random" distribution
df_filtered <- df_replicates %>%
  filter(distribution != "random" & complete.cases(mean_v_evolr) & !is.na(distribution))

# Update the combined dataset for the last plot
df_combined <- df_filtered %>%
  select(distribution, mean_e_fcij, mean_v_fcij) %>%
  pivot_longer(
    cols = c(mean_e_fcij, mean_v_fcij),
    names_to = "group",
    values_to = "value"
  ) %>%
  mutate(group = recode(group, 
                        mean_e_fcij = "Exploiters", 
                        mean_v_fcij = "Victims"))

# Recreate the plots without "random"
dist_er_e <- ggplot(df_filtered, aes(x = distribution, y = mean_e_evolr, fill = distribution)) +
  geom_violin() +
  labs(x = NULL, y = expression(paste("Evolutionary rate ", italic((sigma))^2)), fill = "Distribution") +
  scale_fill_manual(values = c("#E69F00", "#E69F00")) +
  theme_classic() +
  my_theme +
  theme(legend.position = "none")

dist_er_v <- ggplot(df_filtered, aes(x = distribution, y = mean_v_evolr, fill = distribution)) +
  geom_violin() +
  labs(x = NULL, y = expression(paste("Evolutionary rate ", italic((sigma))^2)), fill = "Distribution") +
  scale_fill_manual(values = c("#009E73", "#009E73")) +
  theme_classic() +
  my_theme +
  theme(legend.position = "none")

dist_s_e <- ggplot(df_filtered, aes(x = distribution, y = mean_e_fcij, fill = distribution)) +
  geom_violin() +
  labs(x = NULL, y = expression(paste("Trait matching fluctuation ", (italic(lambda)))), fill = "Distribution") +
  scale_fill_manual(values = c("#E69F00", "#E69F00")) +
  theme_classic() +
  my_theme +
  theme(legend.position = "none")

dist_s_v <- ggplot(df_filtered, aes(x = distribution, y = mean_v_fcij, fill = distribution)) +
  geom_violin() +
  labs(x = NULL, y = expression(paste("Trait matching fluctuation ", (italic(lambda)))), fill = "Distribution") +
  scale_fill_manual(values = c("#009E73", "#009E73")) +
  theme_classic() +
  my_theme +
  theme(legend.position = "none")

dist_s_combined <- ggplot(df_combined, aes(x = distribution, y = value, fill = group)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  labs(
    x = NULL,
    y = expression(paste("Trait matching fluctuation ", (italic(lambda)))),
    fill = "Group"
  ) +
  scale_fill_manual(values = c("Exploiters" = "#E69F00", "Victims" = "#009E73")) +
  theme_classic() +
  my_theme

# Arrange plots
dist_er_matching <- ggarrange(
  dist_er_e, dist_er_v,
  dist_s_e, dist_s_v,
  labels = LETTERS[13:16],
  ncol = 4,
  nrow = 1
)




```


arranged plot evol and matching ~ structure

```{r}

# Wrap individual plots with legends on top
panel_modnest_evolr_e <- ggarrange(plot_modnest_evolr_e, ncol = 1, nrow = 1)
panel_modnest_evolr_v <- ggarrange(plot_modnest_evolr_v, ncol = 1, nrow = 1)
panel_modnest_fcij_combined <- ggarrange(plot_modnest_fcij_combined, ncol = 1, nrow = 1)

plot_er_s_str <- ggarrange(
  s_er_e, s_er_v, s_s_combined,
  c_er_e, c_er_v, c_s_combined,
  panel_modnest_evolr_e, panel_modnest_evolr_v, panel_modnest_fcij_combined,
  dist_er_e, dist_er_v, dist_s_combined,
  ncol = 3,
  nrow = 4,
  common.legend = TRUE,
  labels = LETTERS[1:12]
)
plot_er_s_str

#ggsave(here::here("figures/fig_2_22june2025.png"), height = 14, width = 12)

```




# effects ~ structure


```{r}
df_replicates = df_replicates %>% 
mutate(mean_evenness = sapply(evenness, mean))

df_plot <- df_replicates %>% 
  filter(test_metric == "nestedness") %>% 
  select(nestedness, 
         mean_evenness,
         NODF,
         modularity) %>%
  group_by(nestedness) 


plot_nest_H = ggplot(data = df_plot, aes(x = NODF, y = mean_evenness)) +
  geom_point(alpha = 0.03, size = 3) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Nestedness", y = expression(paste("Biotic effects evennness ",italic((H[b]))))) +
  theme_classic()+
  my_theme


plot_mod_H = ggplot(data = df_plot, aes(x = modularity, y = mean_evenness)) +
  geom_point(alpha = 0.03, size = 3) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Modularity", y = expression(paste("Biotic effects evennness ",italic((H[b]))))) +
  theme_classic()+
  my_theme


df_plot <- df_replicates %>% 
  filter(test_metric == "connectance") %>% 
  select(connectance, 
         mean_evenness) %>%
  group_by(connectance) 


plot_c_H = ggplot(data = df_plot, aes(x = connectance, y = mean_evenness)) +
  geom_point(alpha = 0.03, size = 3) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Connectance", y =expression(paste("Biotic effects evennness ",italic((H[b]))))) +
  theme_classic()+
  my_theme



df_plot <- df_replicates %>%
  filter(test_metric == "distribution", distribution != "random") %>% 
  select(connectance, distribution, mean_evenness) %>%
  group_by(distribution)

# Create the updated plot_dist_H without "random"
plot_dist_H <- ggplot(df_plot, aes(x = distribution, y = mean_evenness)) +
  geom_violin(fill = "gray") +
  labs(x = NULL, y = expression(paste("Biotic effects evenness ", italic((H[b]))))) +
  theme_classic() +
  theme(legend.position = "none") +
  my_theme




df_replicates <- df_replicates %>%
  mutate(degree = map2(degree_v, degree_e, c))

df_plot = data.frame("degree" = unlist(df_replicates$degree),
                     "evenness" = unlist(df_replicates$evenness))

plot_degree_H = ggplot(data = df_plot, aes(x = degree, y = evenness)) +
  geom_point(alpha = 0.008, size = 1) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species degree", y = expression(paste("Biotic effects evennness ",italic((H[b]))))) +
  theme_classic()+
  my_theme


evenness_plots = ggarrange(
  
  plot_c_H,
  plot_mod_H,
  plot_dist_H,
  plot_degree_H,
  
  ncol = 2,
  nrow = 2,
  
  labels = LETTERS[1:4]
  
)

evenness_plots

#ggsave("../figures/supplmat/evenness_str_simulated_corrected.png", height = 8, width = 8)

```


```{r}
df_replicates_random <- df_replicates[df_replicates$distribution == "random",] %>%
  mutate(degree = map2(degree_v, degree_e, c))

df_plot = data.frame("degree" = unlist(df_replicates_random$degree),
                     "evenness" = unlist(df_replicates_random$evenness))

plot_degree_H_random = ggplot(data = df_plot, aes(x = degree, y = evenness)) +
  geom_point(alpha = 0.008, size = 1) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species degree", y = expression(paste("Biotic effects evennness ",italic((H[b]))))) +
  theme_classic()+
  my_theme



df_replicates_sf <- df_replicates[df_replicates$distribution == "scale-free",] %>%
  mutate(degree = map2(degree_v, degree_e, c))

df_plot = data.frame("degree" = unlist(df_replicates_sf$degree),
                     "evenness" = unlist(df_replicates_sf$evenness))

plot_degree_H_sf = ggplot(data = df_plot, aes(x = degree, y = evenness)) +
  geom_point(alpha = 0.008, size = 1) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species degree", y = expression(paste("Biotic effects evennness ",italic((H[b]))))) +
  theme_classic()+
  my_theme



df_replicates_nest <- df_replicates[df_replicates$distribution == "nested",] %>%
  mutate(degree = map2(degree_v, degree_e, c))

df_plot = data.frame("degree" = unlist(df_replicates_nest$degree),
                     "evenness" = unlist(df_replicates_nest$evenness))

plot_degree_H_nest = ggplot(data = df_plot, aes(x = degree, y = evenness)) +
  geom_point(alpha = 0.008, size = 1) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species degree", y = expression(paste("Biotic effects evennness ",italic((H[b]))))) +
  theme_classic()+
  my_theme


plot_degree_h_dist = ggarrange(
  
  plot_degree_H_random,
  plot_degree_H_sf,
  plot_degree_H_nest,
  
  labels = LETTERS[1:3],
  
  ncol = 3,
  nrow = 1
  
)

plot_degree_h_dist

#ggsave("../figures/supplmat/plot_degree_h_dist.png", height = 5, width = 12)

```




# FIGURE 1


```{r}
plot_complexity_effects <- readRDS(here::here("figures/plot_complexity_effects.rds"))
p_mod_nest_effects <- readRDS(here::here("figures/p_mod_nest_effects.rds"))
plot_degree_effects <- readRDS(here::here("figures/plot_degree_effects.rds"))


fig_1 <- ggarrange(
  
  plot_c_H, 
  plot_mod_H, 
  plot_dist_H,
  plot_complexity_effects,
  p_mod_nest_effects,
  plot_degree_effects,
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  labels = LETTERS[1:6],
  legend = "bottom")

fig_1

#ggsave("../figures/fig_1.png", height = 9, width = 11)
```



# adaptation ~ structure


```{r}

# RICHNESS


df_plot <- df_replicates %>% 
  filter(test_metric == "richness") %>% 
  select(connectance, n,
         mean_e_load_abiot,
         mean_e_load_biot,
         mean_e_load_network,
         mean_v_load_abiot,
         mean_v_load_biot,
         mean_v_load_network,
         NODF,
         modularity) %>%
  group_by(n) %>%
  summarize(mean_mean_e_load_abiot = mean(mean_e_load_abiot),
            mean_mean_e_load_biot = mean(mean_e_load_biot),
            mean_mean_e_load_network = mean(mean_e_load_network),
            mean_mean_v_load_abiot = mean(mean_v_load_abiot),
            mean_mean_v_load_biot = mean(mean_v_load_biot),
            mean_mean_v_load_network = mean(mean_v_load_network),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])


s_adapt_e = ggplot(data = df_plot, aes(x = mean_mean_e_load_abiot, y = mean_mean_e_load_biot)) +
  geom_point(aes(color = n), alpha = 0.5, size = 4) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
scale_color_viridis_c(name = "Species richness", option = "turbo") +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
  theme(legend.position = "top",
        legend.key.size = unit(0.5, "cm"))+
    my_theme


s_adapt_v = ggplot(data = df_plot, aes(x = mean_mean_v_load_abiot, y = mean_mean_v_load_biot)) +
  geom_point(aes(color = n), alpha = 0.5, size = 4, shape = "triangle") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
scale_color_viridis_c(name = "Species richness", option = "turbo") +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
  theme(legend.position = "top",
        legend.key.size = unit(0.5, "cm"))+
    my_theme


# CONNECTANCE


df_plot <- df_replicates %>% 
  filter(test_metric == "connectance") %>% 
  select(connectance, n,
         mean_e_load_abiot,
         mean_e_load_biot,
         mean_e_load_network,
         mean_v_load_abiot,
         mean_v_load_biot,
         mean_v_load_network,
         NODF,
         modularity) %>%
  group_by(connectance) %>%
  summarize(mean_mean_e_load_abiot = mean(mean_e_load_abiot),
            mean_mean_e_load_biot = mean(mean_e_load_biot),
            mean_mean_e_load_network = mean(mean_e_load_network),
            mean_mean_v_load_abiot = mean(mean_v_load_abiot),
            mean_mean_v_load_biot = mean(mean_v_load_biot),
            mean_mean_v_load_network = mean(mean_v_load_network),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])


c_adapt_e = ggplot(data = df_plot, aes(x = mean_mean_e_load_abiot, y = mean_mean_e_load_biot)) +
  geom_point(aes(color = connectance), alpha = 0.5, size = 4) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
scale_color_viridis_c(name = expression(italic(C)), option = "turbo") +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
  theme(legend.position = "top",
        legend.key.size = unit(0.5, "cm"))+
    my_theme


c_adapt_v = ggplot(data = df_plot, aes(x = mean_mean_v_load_abiot, y = mean_mean_v_load_biot)) +
  geom_point(aes(color = connectance), alpha = 0.5, size = 4, shape = "triangle") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
scale_color_viridis_c(name = expression(italic(C)), option = "turbo") +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
  theme(legend.position = "top",
        legend.key.size = unit(0.5, "cm"))+
    my_theme



# MODULARITY


df_plot <- df_replicates %>% 
  filter(test_metric == "nestedness") %>% 
  select(nestedness,connectance, n,
         mean_e_load_abiot,
         mean_e_load_biot,
         mean_e_load_network,
         mean_v_load_abiot,
         mean_v_load_biot,
         mean_v_load_network,
         NODF,
         modularity) %>%
  group_by(nestedness) %>%
  summarize(mean_mean_e_load_abiot = mean(mean_e_load_abiot),
            mean_mean_e_load_biot = mean(mean_e_load_biot),
            mean_mean_e_load_network = mean(mean_e_load_network),
            mean_mean_v_load_abiot = mean(mean_v_load_abiot),
            mean_mean_v_load_biot = mean(mean_v_load_biot),
            mean_mean_v_load_network = mean(mean_v_load_network),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])


mod_adapt_e = ggplot(data = df_plot, aes(x = mean_mean_e_load_abiot, y = mean_mean_e_load_biot)) +
  geom_point(aes(color = modularity), alpha = 0.5, size = 4) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
scale_color_viridis_c(name = "modularity", option = "turbo") +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
  theme(legend.position = "top",
        legend.key.size = unit(0.5, "cm"))+
    my_theme


mod_adapt_v = ggplot(data = df_plot, aes(x = mean_mean_v_load_abiot, y = mean_mean_v_load_biot)) +
  geom_point(aes(color = modularity), alpha = 0.5, size = 4, shape = "triangle") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
scale_color_viridis_c(name = "modularity", option = "turbo") +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
  theme(legend.position = "top",
        legend.key.size = unit(0.5, "cm"))+
    my_theme



# DSITRIBUTION


df_plot <- df_replicates %>% 
  filter(test_metric == "distribution") %>% 
  select(distribution,nestedness,connectance, n,
         mean_e_load_abiot,
         mean_e_load_biot,
         mean_e_load_network,
         mean_v_load_abiot,
         mean_v_load_biot,
         mean_v_load_network,
         NODF,
         modularity) %>%
  group_by(distribution) %>%
  summarize(mean_mean_e_load_abiot = mean(mean_e_load_abiot),
            mean_mean_e_load_biot = mean(mean_e_load_biot),
            mean_mean_e_load_network = mean(mean_e_load_network),
            mean_mean_v_load_abiot = mean(mean_v_load_abiot),
            mean_mean_v_load_biot = mean(mean_v_load_biot),
            mean_mean_v_load_network = mean(mean_v_load_network),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])


df_filtered <- df_replicates[complete.cases(df_replicates$mean_v_evolr) & !is.na(df_replicates$distribution), ]


dist_adapt_e = ggplot(data = df_filtered, aes(x = mean_e_load_abiot, y = mean_e_load_biot)) +
  geom_point(aes(color = distribution), alpha = 0.1, size = 4) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
scale_color_manual(values = c("gray", "#0072B2", "#CC79A7")) +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
    my_theme


dist_adapt_v = ggplot(data = df_filtered, aes(x = mean_v_load_abiot, y = mean_v_load_biot)) +
  geom_point(aes(color = distribution), alpha = 0.1, size = 4, shape = "triangle") +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(paste("abiotic load ", italic((L[alpha])))), y = expression(paste("biotic load ", italic((L[beta]))))) +
scale_color_manual(values = c("gray", "#0072B2", "#CC79A7")) +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
    my_theme




s_adapt_str = ggarrange(
  
  s_adapt_e, s_adapt_v,
  
  ncol = 2,
  nrow = 1,
  
  common.legend = TRUE,
  
    legend = "right",
  
  labels = LETTERS[1:2]
)



c_adapt_str = ggarrange(
  
  c_adapt_e, c_adapt_v,
  
  ncol = 2,
  nrow = 1,
  
  common.legend = TRUE,
  
    legend = "right",
  
  labels = LETTERS[3:4]
)

mod_adapt_str = ggarrange(
  
  mod_adapt_e, mod_adapt_v,
  
  ncol = 2,
  nrow = 1,
  
  common.legend = TRUE,
  
    legend = "right",
  
  labels = LETTERS[5:6]
)


dist_adapt_str = ggarrange(
  
  dist_adapt_e, dist_adapt_v,
  
  ncol = 2,
  nrow = 1,
  
  common.legend = TRUE,
  
  legend = "right",
  
  labels = LETTERS[7:8]
)




plot_adapt_str = ggarrange(
 
  s_adapt_str,
  c_adapt_str,
  mod_adapt_str,
  dist_adapt_str,
  
  ncol = 1,
  nrow = 4
  
  
)

plot_adapt_str

#ggsave("../figures/supplmat/adapt_str_simulated_corrected.png", height = 12, width = 9)


```



Network load:

```{r}

# RICHNESS

df_plot <- df_replicates %>% 
  filter(test_metric == "richness") %>% 
  select(connectance, n,
         mean_comm_load_network,
         NODF,
         modularity) %>%
  group_by(n) %>%
  summarize(mean_mean_comm_load_network = mean(mean_comm_load_network),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])

s_adapt3rd = ggplot(data = df_plot, aes(x = n, y = mean_mean_comm_load_network)) +
  geom_point(alpha = 0.5, size = 4) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "Species richness", y = expression(paste("network load ", italic((L[gamma]))))) +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
    my_theme


# CONNECTANCE

df_plot <- df_replicates %>% 
  filter(test_metric == "connectance") %>% 
  select(connectance, n,
         mean_comm_load_network,
         NODF,
         modularity) %>%
  group_by(connectance) %>%
  summarize(mean_mean_comm_load_network = mean(mean_comm_load_network),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])

c_adapt3rd = ggplot(data = df_plot, aes(x = connectance, y = mean_mean_comm_load_network)) +
  geom_point(alpha = 0.5, size = 4) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = expression(italic(C)), y = expression(paste("network load ", italic((L[gamma]))))) +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
    my_theme


# MODULARITY

df_plot <- df_replicates %>% 
  filter(test_metric == "nestedness") %>% 
  select(connectance, n,
         mean_comm_load_network,
         NODF,
         modularity) %>%
  group_by(modularity) %>%
  summarize(mean_mean_comm_load_network = mean(mean_comm_load_network),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])

mod_adapt3rd = ggplot(data = df_plot, aes(x = modularity, y = mean_mean_comm_load_network)) +
  geom_point(alpha = 0.5, size = 4) +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "modularity", y = expression(paste("network load ", italic((L[gamma]))))) +
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
    my_theme



# DISTRIBUTION

df_plot <- df_replicates %>% 
  filter(test_metric == "distribution") %>% 
  select(distribution,connectance, n,
         mean_comm_load_network,
         NODF,
         modularity) %>%
  group_by(distribution) %>%
  summarize(mean_mean_comm_load_network = mean(mean_comm_load_network),
            modularity = modularity[1],
            NODF = NODF[1],
            connectance[1])

dist_adapt3rd = ggplot(data = df_filtered, aes(x = distribution, y = mean_comm_load_network)) +
  geom_violin() +
# geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = NULL, y = expression(paste("network load ", italic((L[gamma]))))) +
   scale_fill_manual(values = c("gray", "gray30", "black"))+
  scale_y_continuous(labels = back_transform_sqrt)+
  theme_classic()+
    my_theme


p_3rdterm_str = ggarrange(
  
  
  s_adapt3rd,
  c_adapt3rd,
  mod_adapt3rd,
  dist_adapt3rd,
  
  ncol = 2,
  nrow = 2,
  
  labels = LETTERS[1:4]
  
  
)

p_3rdterm_str

#ggsave("../figures/supplmat/3rdterm_simstr_corrected.png", height = 8, width = 8)


```

