---
title: "Demonstration: compute total effects on species in a network"
---

This script is a demonstration of how we compute the net or total effects of interactions (direct + indirect) on each species of a network, and the mean and sd of the community.



```{r}
library(bipartite)
library(igraph)
library(reshape2)
library(tidyverse)
library(ggplot2)
library(ggpubr)
```



Functions to generate a network given connectance, where all species have at least one interaction (just a way to generate random networks):

```{r}
# function to generate probabilities of interaction based on trait matching

generate_prob_matrix_each_sim <- function(n.host, n.exploiter, trait_vec, scale = 0.1) {
  
  nsp = n.host + n.exploiter
  dif.traits <- outer(trait_vec, trait_vec, FUN= "-")
  
  # Create trait matching matrix
  
  matching_matrix<-dif.traits[(n.exploiter+1):nsp, 1:n.exploiter]
  
  # compute probabilities based on trait matching
  
  prob_matching<-exp(-(matching_matrix/scale)^2)
  
  return (prob_matching)
  
}



# Function to generate network given connectance


generate_network_given_conn_each_sim_allsp.int <- function(n_exploiter, n_host, conn, trait_sim) {
  
  prob_matrix<-generate_prob_matrix_each_sim(n.host = n_host, n.exploiter = n_exploiter, trait_vec = trait_sim)
  
  #n_ones <- round(n_exploiter*n_host*conn)
  
  q_prob_int<- quantile(prob_matrix, prob = 1 - conn)
  index<-which(prob_matrix > q_prob_int, arr.ind = TRUE)
  L <- matrix(0, nr = n_host, nc = n_exploiter)
  L[index] <-1
  
  # Add one interaction to exploiters with no interactions
  
  expl_no_int<-colSums(L)
  vec_e_no_int<-which(expl_no_int == 0)
  
  for (i in vec_e_no_int) {
    v_toeat <- which.max(prob_matrix[,i])
    L[v_toeat,i] <- 1
  }
  
  
  # Add one interaction to hosts with no interactions
  
  host_no_int<-rowSums(L)
  vec_h_no_int<-which(host_no_int == 0)
  
  for (i in vec_h_no_int) {
    e_predating <- which.max(prob_matrix[i,])
    L[i,e_predating] <- 1
  }
  
  return(L)
  
}


# compute F matrix from binary matrix

create_F <- function(binmat) {
  
  mat_F<-rbind (
    cbind( matrix(0, nrow=nrow(binmat), ncol=nrow(binmat)),binmat),
    cbind( t(binmat), matrix(0, nrow=ncol(binmat),ncol=ncol(binmat)))) 
  return(mat_F)
}

```



Functions to compute the total effects (direct and indirect) that species experience in a network:

```{r}

compute_total_effects <- function(int_matrix) {
  
  f_mat <- create_F(int_matrix)
  
  # generating a matrix that all rows sum to one. In this matrix the element describes the proportional effect of a species in the column on the species in the row
  
  P = f_mat
  
  for (i in 1:nrow(f_mat)) {
    P[i,] <- P[i,] / sum(P[i,])
  }
  
  #weighting each row for the total contribution of interactions for a given process (e.g., the amount of selection imposed by mutualistic partners on the mean trait of a given species
  
  m=0.6 # m is the strength and needs to be between zero and one
  Q=m*P 
  
  #Note: you can change m for a diagonal matrix if you want m to be species specific
  
  #computing the matrix of total effects (direct and indirect)
  
  I <- diag(nrow(f_mat)) #identity matrix
  T <- solve(I - Q)  #compute the inverse of I-Q
  
  return(T)
  
}

```


Function to calculate Shannon Index for a row (i.e. total effects that each species receives):


```{r}
calculate_shannon <- function(row) {
  
  eps <- 1e-10  # Small constant to avoid taking the logarithm of zero
  row_proportions <- row / sum(row)
  row_proportions[row_proportions == 0] <- eps  # Replace zeros with eps
  shannon_index <- -sum(row_proportions * log2(row_proportions))
  normalized_shannon <- shannon_index / log2(length(row))
  return(normalized_shannon)
}

```



Function to compute total effects on each species, mean and sd total effects of a network, and save objects (binary matrix, T matrix, vector of effects, network plot):

```{r}

# Function to compute the mean total effects that species experience in a network


compute_df_effects <- function(n.exploiter, n.host, conn, seed, id) {
  
  label <- id
  set.seed(seed)
  
  # create network
  
  nsp <- n.exploiter + n.host
  vec_traits <- rnorm(nsp, mean=0, sd=0.1)
  
  binary.matrix <- generate_network_given_conn_each_sim_allsp.int (
    n_exploiter = n.exploiter, n_host = n.host, 
    conn = conn,
    trait_sim = vec_traits)
  
  # record network properties
  
  size <- nsp
  c <- networklevel(binary.matrix, index = "connectance")
  mod <- networklevel(binary.matrix, index = "modularity")
  NODF <- networklevel(binary.matrix, index = "NODF")
  
  # compute T
  
  T_binary.matrix <- compute_total_effects(binary.matrix)
  T_mean <- mean(T_binary.matrix)
  
  # compute mean and sd spp T inn (received effects)
  
  ### Calculate normalized Shannon Index for each row
  normalized_shannon_values <- numeric(nrow(T_binary.matrix))
  
  for (i in 1:nrow(T_binary.matrix)) {
    normalized_shannon_values[i] <- calculate_shannon(T_binary.matrix[i, ])
  }
  
  T_inn_vec <- normalized_shannon_values
  mean_T_spp_inn <- mean(normalized_shannon_values)
  sd_T_spp_inn <- sd(normalized_shannon_values)
  
  # compute mean and sd spp T out
  
  T_out_vec <- colSums(T_binary.matrix)
  mean_T_spp_out <- mean(colSums(T_binary.matrix))
  sd_T_spp_out <- sd(colSums(T_binary.matrix))
  
  # PLOT - network and total effects received by each node
  
  # create graph object
  
  graph <- graph_from_incidence_matrix(binary.matrix, weighted = NULL, directed = FALSE)
  
  # compute degree
  
  d_vec <- degree(graph)
  
  V(graph)$degree <- d_vec
  V(graph)$normalized_shannon_values <- normalized_shannon_values
  
  df <- data.frame(
    
    label,
    size,
    c,
    mod,
    NODF,
    
    T_mean,
    
    mean_T_spp_inn,
    sd_T_spp_inn,
    
    mean_T_spp_out,
    sd_T_spp_out
    
    )
  
  
  l_graph <- list()
  l_graph[[1]] <- graph
  df$graph <- l_graph
  
  l_vec_traits <- list()
  l_vec_traits[[1]] <- graph
  df$vec_traits <- l_vec_traits
  
  l_T_binary.matrix <- list()
  l_T_binary.matrix[[1]] <- T_binary.matrix
  df$T_binary.matrix <- l_T_binary.matrix
  
  l_binary.matrix <- list()
  l_binary.matrix[[1]] <- binary.matrix
  df$binary.matrix <- l_binary.matrix
  
  l_T_inn_vec <- list()
  l_T_inn_vec[[1]] <- T_inn_vec
  df$T_inn_vec <- l_T_inn_vec
  
  l_T_out_vec <- list()
  l_T_out_vec[[1]] <- T_out_vec
  df$T_out_vec <- l_T_out_vec
  
  return(df)
  
}


```



Compute the total effects for networks varying in size and connectance

```{r}

df_effects_A <- compute_df_effects(n.exploiter = 8, 
                   n.host = 8, 
                   conn = 0.2, 
                   seed = 100, 
                   id = "A")

df_effects_B <- compute_df_effects(n.exploiter = 25, 
                   n.host = 25, 
                   conn = 0.2, 
                   seed = 100, 
                   id = "B")

df_effects <- rbind(df_effects_A, df_effects_B)

df_effects
```



# Plots - T matrix

Function to plot T matrix as heatmap

```{r}

plot_heatmap <- function(T_matrix, nsp) {
  

# the matrix needs names
dat <- T_matrix
rownames(dat) <- 1:nsp
colnames(dat) <- 1:nsp

## convert to tibble, add row identifier, and shape "long"

dat2 <- dat %>%
  as_tibble() %>%
  rownames_to_column("Var1") %>%
  pivot_longer(-Var1, names_to = "Var2", values_to = "value") %>%
  mutate(
    Var1 = factor(Var1, levels = 1:nsp),
    Var2 = factor(gsub("V", "", Var2), levels = 1:nsp)
  )


plot <- ggplot(dat2, aes(Var1, Var2)) +
  geom_tile(aes(fill = value)) +
 # geom_text(aes(label = round(value, 1))) +
  scale_fill_gradient(low = "white", high = "red")+
  theme_classic()

return(plot)
  
}

```


*Plot T matrices as heatmaps*

```{r}

p_heatmap_a <- plot_heatmap(df_effects_A$T_binary.matrix[[1]], nsp = 16)
p_heatmap_b <- plot_heatmap(df_effects_B$T_binary.matrix[[1]], nsp = 50)

ggarrange(p_heatmap_a,
          p_heatmap_b,
          labels = c("A", "B"))

```


*Plots - network and total effects on each spp*

Lighter colors indicates higher values.

```{r}

# Define a color scale based on the range of normalized_shannon_values
color_scale <- heat.colors(100)

graph <- df_effects_A$graph[[1]]

plot(
  graph,
  vertex.size = V(graph)$degree * 6,
  vertex.color = color_scale[cut(V(graph)$normalized_shannon_values, breaks = 100)],
  layout = layout_as_bipartite,
  vertex.label = NA,
  main = "A"
)

graph <- df_effects_B$graph[[1]]

plot(
  graph,
  vertex.size = V(graph)$degree * 2,
  vertex.color = color_scale[cut(V(graph)$normalized_shannon_values, breaks = 100)],
  layout = layout_as_bipartite,
  vertex.label = NA,
  main = "B"
)

```



